# Descriptive analyses in srvyr {#c06}

## Goals



## Introduction

Descriptive analysis allows you to investigate your data and gain insight into the data. Common descriptive analyses include calculating mean, median of numeric data or proportions in categorical data.



      
## Overview of descriptive analysis using `srvyr` package

1. Create a `tbl_svy` object using `srvyr::as_survey_design()` or `srvyr::as_survey_rep()`
2. Subset the data for subpopulations, if needed, using `dplyr::filter()`
3. Specify domains of analysis using `dplyr::group_by()` 
4. Within `srvyr::summarize()`, specify variables to calculate means, totals, proportions, quantiles, and more

### A brief refresher on the `dplyr::summarize()` function

The `dplyr::summarize()` function collapses many values down to a single summary:

<!-- TODO: Add example -->

These verbs can be used in conjunction with `group_by()`, applying the functions on a group-by-group basis to create grouped summaries.
  
## Categorical data

_Categorical data_, or the [definition], 

Analyzing categorical data lets us...

Common analysis for categorical data include:

* Weighted proportions
* Weighted counts
* Unweighted proportions
* Unweighted counts

With the ACES data, we create a survey object (as specified in Chapter 05):

Create a `tbl_svy` object using: `as_survey_design` or `as_survey_rep`

```{r anes_des, error=FALSE, warning=FALSE}
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

anes <- read_rds(here("Data", "anes_2020.rds")) %>%
  mutate(Weight = Weight / sum(Weight) * 231592693) 
# adjust weight to sum to citizen pop, 18+ in Nov 2020 per ANES methodology documentation

anes_des <- anes %>%
  as_survey_design(
    weights = Weight,
    strata = Stratum,
    ids = VarUnit,
    nest = TRUE
  )
```

### `survey_count()` Syntax

The `srvyr::survey_count()` syntax is very similar to the `dplyr::count()` syntax. It can only be called on `tbl_srvy()` objects.

Similarly, `srvyr::count()` can take one or many variables.

```{r}

```


We don’t need a `wt` argument because we defined the weights when we set up the `tbl_srvy()` object.


- `wt` adds additional weights in addition to the survey weights, by default is `NULL`
- you can sort the variables, by default is `FALSE`
- you can edit the name of the variable created to make it something other than n
- you can change the variance type with `vartype`, defaults to standard error

- `survey_count` functions similarly to `count` in that it is <b>NOT</b> called within `summarize`
- Produces weighted counts and variance of your choice of those counts
```{r survey_count_syn, eval=FALSE, tidy=FALSE}
survey_count(
   x,
   ...,
   wt = NULL,
   sort = FALSE,
   name = "n",
   .drop = dplyr::group_by_drop_default(x),
   vartype = c("se", "ci", "var", "cv")
)
```


### Weighted proportions

- Functions to use within `summarize` after `group_by`
  - survey_mean/survey_prop
  - survey_total

The `srvyr::survey_mean()` calculates the means and portions from complex survey data. The function wraps around `survey::svymean()` (if `proportion - FALSE`) and `survey::svyciprop()` (if `proportion = TRUE`).



