# Statistical testing {#c06-statistical-testing}

```{r}
#| label: stattest-styler
#| include: false
knitr::opts_chunk$set(tidy = 'styler')
```


::: {.prereqbox-header}
`r if (knitr:::is_html_output()) '### Prerequisites {- #prereq6}'`
:::

::: {.prereqbox data-latex="{Prerequisites}"}
For this chapter, load the following packages:
```{r}
#| label: stattest-setup
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(survey) 
library(srvyr) 
library(srvyrexploR)
library(broom)
library(gt)
library(prettyunits)
```

We are using data from ANES and RECS described in Chapter \@ref(c04-getting-started). As a reminder, here is the code to create the design objects for each to use throughout this chapter. For ANES, we need to adjust the weight so it sums to the population instead of the sample (see the ANES documentation and Chapter \@ref(c04-getting-started) for more information).

```{r}
#| label: stattest-anes-des
#| eval: FALSE
targetpop <- 231592693
data(anes_2020)

anes_adjwgt <- anes_2020 %>%
  mutate(Weight = Weight / sum(Weight) * targetpop)

anes_des <- anes_adjwgt %>%
  as_survey_design(
    weights = Weight,
    strata = Stratum,
    ids = VarUnit,
    nest = TRUE
  )
```

For RECS, details are included in the RECS documentation and Chapters \@ref(c04-getting-started) and \@ref(c10-specifying-sample-designs).

```{r}
#| label: stattest-recs-des
#| eval: FALSE
data(recs_2020)

recs_des <- recs_2020 %>%
  as_survey_rep(
    weights = NWEIGHT,
    repweights = NWEIGHT1:NWEIGHT60,
    type = "JK1",
    scale = 59/60,
    mse = TRUE
  )
```
:::

## Introduction

When analyzing results from a survey, the point estimates described in Chapter \@ref(c05-descriptive-analysis) help us understand the data at a high level. Still, researchers and the public often want to make comparisons between different groups. These comparisons are calculated through statistical testing. 

The general idea of statistical testing is the same for data obtained through surveys and data obtained through other methods, where we compare the point estimates and variance estimates of each statistic to see if statistically significant differences exist. However, statistical testing for complex surveys involves additional considerations due to the need to account for the sampling design in order to obtain accurate variance estimates.

Statistical testing, also called hypothesis testing, involves declaring a null and alternative hypothesis. A null hypothesis is denoted as $H_0$ and the alternative hypothesis is denoted as $H_A$. The null hypothesis is the default assumption in that there are no differences in the data, or that the data are operating under "standard" behaviors.  On the other hand, the alternative hypothesis is the break from the "standard" and what we are trying to determine if the data support. 

Let's review an example outside of survey data. If we are flipping a coin, a null hypothesis would be that the coin is fair and that each side has an equal chance of being flipped.  In other words, the probability of the coin landing on each side is 1/2.  Whereas an alternative hypothesis could be that the coin is unfair and that one side has a higher probability of being flipped (e.g., a probability of 1/4 to get heads, but a probability of 3/4 to get tails). We write this set of hypotheses as:

- $H_0: \rho_{heads} = \rho_{tails}$, where $\rho_{x}$ is the probability of flipping the coin and having it land on heads ($\rho_{heads}$) or tails ($\rho_{tails}$)
- $H_A: \rho_{heads} \neq \rho_{tails}$

When we conduct hypothesis testing, the statistical models calculate a p-value, which shows how likely we are to observe the data if the null hypothesis is true. If the p-value (a probability between 0 and 1) is small, we have strong evidence to reject the null hypothesis as it is unlikely to see the data we are observing if the null hypothesis is true. However, if the p-value is large, we say we do not have evidence to reject the null hypothesis. The size of the p-value for this cut off is determined by type 1 error known as $\alpha$.  A common type 1 error value for statistical testing is to use $\alpha = 0.05$.^[For more information on statistical testing, we recommend reviewing introduction to statistics textbooks.] It is common for explanations of statistical testing to refer to confidence level.  The confidence level is the inverse of the type 1 error.  Thus, if $\alpha = 0.05$, the confidence level would be 95%.

The functions in the {survey} package allow for the correct estimation of the variances. This chapter covers the following statistical tests with survey data and functions:

* Comparison of proportions `svyttest()`
* Comparison of means `svyttest()`
* Goodness of fit tests `svygofchisq()`
* Tests of independence `svychisq()`
* Tests of homogeneity `svychisq()`

## Dot Notation {#dot-notation}

Up to this point, we have shown functions that use wrappers from the {srvyr} package. This means that the functions work with tidyverse syntax. However, the functions in this chapter do not have wrappers in the {srvyr} package and are instead used directly from the {survey} package. Therefore, the design object is *not* the first argument, and to use these functions with the magrittr pipe (`%>%`) and tidyverse syntax, we need to use dot (`.`) notation^[This could change in the future if another package is built or {srvyr} is expanded to work with **tidymodels** but no such plans are known at this time.]

Functions that work with the magrittr pipe (`%>%`) have the dataset as the first argument. When we run a function with the pipe, it automatically places anything to the left of the pipe into the first argument of the function to the right of the pipe. For example, if we wanted to take the `mtcars` data and filter to cars with six cylinders, we can write the code in at least four different ways:

1. `filter(mtcars, cyl == 6)`
2. `mtcars %>% filter(cyl == 6)`
3. `mtcars %>% filter(., cyl == 6)`
4. `mtcars %>% filter(.data = ., cyl == 6)`

Each of these lines of code produces the same output since the argument that takes the dataset is in the first spot in `filter()`. The first two are probably familiar to those who have worked with the tidyverse. The third option functions the same way as the second one but is explicit that `mtcars` goes into the first argument, and the fourth option indicates that `mtcars` is going into the named argument of `.data`. Here, we are telling R to take what's on the left side of the pipe (`mtcars`) and pipe it into the spot with the dot (`.`)---the first argument.

In functions that are not part of the tidyverse, the data argument may not be in the first spot. For example, in `svyttest()`, the data argument is in the second spot, which means we need to place the dot (`.`) in the second spot and not the first. For example:

```r
svydata_des %>%
 svyttest(x ~ y, .)
```

By default, the pipe places the left-hand object in the first argument spot. Placing the dot (`.`) in the second argument spot indicates that the survey design object `svydata_des` should be used in the second argument and not the first.

Alternatively, named arguments could be used to place the dot first as named arguments can appear at any location, as in the following:

```r
svydata_des %>%
 svyttest(design = ., x ~ y)
```

However, the following code does not work as the `svyttest()` function expects the formula as the first argument when arguments are not named:

```r
svydata_des %>%
 svyttest(., x ~ y)
```

## Comparison of Proportions and Means {#stattest-ttest}

We use t-tests to compare two proportions or means. T-tests allow us to determine if one proportion or mean is statistically different from another. They are commonly used to determine if a single estimate differs from a known value (e.g., 0 or 50%) or to compare two group means (e.g., North versus South). Comparing a single estimate to a known value is called a *one sample t-test*, and we can set up the hypothesis test as follows:  

  - $H_0: \mu = 0$ where $\mu$ is the mean outcome and $0$ is the value we are comparing it to
  - $H_A: \mu \neq 0$

For comparing two estimates, this is called a *two-sample t-test* and we can set up the hypothesis test as follows:

  - $H_0: \mu_1 = \mu_2$ where $\mu_i$ is the mean outcome for group $i$
  - $H_A: \mu_1 \neq \mu_2$

Two sample t-tests can also be *paired* or *unpaired*. If the data come from two different populations (e.g., North versus South), the t-test run is an *unpaired* or *independent samples* t-test. *Paired* t-tests occur when the data come from the same population. This is commonly seen with data from the same population in two different time periods (e.g., before and after an intervention).

The difference between t-tests with non-survey data and survey data is based on the underlying variance estimation difference.  Chapter \@ref(c10-specifying-sample-designs) provides a detailed overview of the math behind the mean and sampling error calculations for various sample designs. The functions in the {survey} package account for these nuances, provided the design object is correctly defined.

### Syntax {#stattest-ttest-syntax}

When we do not have survey data, we can use the `t.test()` function from the {stats} package. This function does not allow for weights or the variance structure that need to be accounted for with survey data. Therefore, we need to use the `svyttest()` function from {survey} when using survey data. Many of the arguments are the same between the two functions, but there are a few key differences:

  - We need to use the survey design object instead of the original data frame
  - We can only use a formula and not separate x and y data
  - The confidence level cannot be specified and is always be set to 95%. However, we show examples of how the confidence level can be changed after running the `svyttest()` function by using the `confint()` function.

Here is the syntax for the `svyttest()` function:

```r
svyttest(formula,
         design,
         ...)
```

The arguments are:

* `formula`: Formula, `outcome~group` for two-sample, `outcome~0` or `outcome~1` for one-sample. The group variable must be a factor or character with two levels, or be coded 0/1 or 1/2. We give more details on formula set-up below for different types of tests.
* `design`: survey design object
* `...`: This passes options on for one-sided tests only, and thus, we can specify `na.rm=TRUE`

Notice that the first argument here is the `formula` and not the `design`. This means we must use the dot `(.)` if we pipe in the survey design object (as described in Section \@ref(dot-notation)).

The `formula` argument can take several different forms depending on what we are measuring. Here are a few common scenarios:

1. **One-sample t-test:** 
    a. **Comparison to 0:** `var ~ 0`, where `var` is the measure of interest, and we compare it to the value `0`. For example, we could test if the population mean of household debt is different from `0` given the sample data collected.
    b. **Comparison to a different value:** `var - value ~ 0`, where `var` is the measure of interest and `value` is what we are comparing to. For example, we could test if the proportion of the population that has blue eyes is different from `25%` by using `var - 0.25 ~ 0`. Note that specifying the formula as `var ~ 0.25` is not equivalent and results in a syntax error.
2. **Two-sample t-test:**
    a. **Unpaired:**
        - **2 level grouping variable:** `var ~ groupVar`, where `var` is the measure of interest and `groupVar` is a variable with two categories. For example, we could test if the average age of the population who voted for president in 2020 differed from the age of people who did not vote. In this case, age would be used for `var`, and a binary variable indicating voting activity would be the `groupVar`.
        - **3+ level grouping variable:** `var ~ groupVar == level`, where `var` is the measure of interest, `groupVar` is the categorical variable, and `level` is the category level to isolate. For example, we could test if the test scores in one classroom differed from all other classrooms where `groupVar` would be the variable holding the values for classroom IDs and `level` is the classroom ID we want to compare to the others.
    b. **Paired:** `var_1 - var_2 ~ 0`, where `var_1` is the first variable of interest and `var_2` is the second variable of interest. For example, we could test if test scores on a subject differed between the start and the end of a course so `var_1` would be the test score at the beginning of the course and `var_2` would be the score at the end of the course.
    
The `na.rm` argument defaults to `FALSE`, which means if any data values are missing, the t-test does not compute. Throughout this chapter, we always set `na.rm = TRUE`, but before analyzing the survey data, review the notes provided in Chapter \@ref(c11-missing-data) to better understand how to handle missing data.

Let's walk through a few examples using the ANES and RECS data.

### Examples {#stattest-ttest-examples}

#### Example 1: One-sample t-test for Mean {.unnumbered #stattest-ttest-ex1}  

RECS asks respondents to indicate what temperature they set their house to during the summer at night.^[During the summer, what is your home’s typical indoor temperature inside your home at night?] In our data, we have called this variable `SummerTempNight`. If we want to see if the average U.S. household sets its temperature at a value different from 68$^\circ$F^[This is the temperature that Stephanie prefers at night during the summer, and she wanted to see if she was different from the population.], we could set up the hypothesis as follows: 

- $H_0: \mu = 68$ where $\mu$ is the average temperature U.S. households set their thermostat to in the summer at night
- $H_A: \mu \neq 68$

To conduct this in R, we use `svyttest()` and subtract the temperature on the left-hand side of the formula:

```{r}
#| label: stattest-ttest-ex1
ttest_ex1 <- recs_des %>%
  svyttest(
    formula = SummerTempNight - 68 ~ 0,
    design = .,
    na.rm = TRUE
  )

ttest_ex1
```

To pull out specific output, we can use R's built-in `$` operator. For instance, to obtain the estimate $\mu - 68$, we run `ttest_ex1$estimate`.

If we want the average, we take our t-test estimate and add it to 68: 

```{r}
#| label: stattest-ttest-ex1-add
ttest_ex1$estimate + 68
```

Or, we can use the `survey_mean()` function described in Chapter \@ref(c05-descriptive-analysis):

```{r}
#| label: stattest-ttest-ex1-svymean
recs_des %>%
  summarize(mu = survey_mean(SummerTempNight, na.rm = TRUE))
```

The result is the same in both methods, so we see that the average temperature U.S. households set their thermostat to in the summer at night is `r signif(ttest_ex1$estimate + 68,3)`$^\circ$F. Looking at the output from `svyttest()`, the t-statistic is `r signif(ttest_ex1$statistic, 3)`, and the p-value is $`r pretty_p_value(ttest_ex1[["p.value"]])`$, indicating that the average is statistically different from 68$^\circ$F at an $\alpha$ level of $0.05$.

If we want an 80% confidence interval for the test statistic, we can use the function `confint()` to change the confidence level. Below, we print both the original 95% confidence interval and the 80% confidence interval:

```{r}
#| label: stattest-ttest-ex1-ci80
confint(ttest_ex1, level = 0.95)
confint(ttest_ex1, level = 0.8)
```

In this case, neither confidence interval contains 0, and we draw the same conclusion from either that the average temperature households set their thermostat in the summer at night is significantly higher than 68$^\circ$F.

#### Example 2: One-sample t-test for Proportion {.unnumbered #stattest-ttest-ex2}

RECS asked respondents if they use any air conditioning (A/C) in their home.^[Is any air conditioning equipment used in your home?] In our data, we call this variable `ACUsed`. Let's look at the proportion of U.S. households that use A/C in their homes using the `survey_prop()` function we learned in Chapter \@ref(c05-descriptive-analysis).

```{r}
#| label: stattest-ttest-acused
acprop <- recs_des %>%
  group_by(ACUsed) %>%
  summarize(p = survey_prop())

acprop
```

Based on this, `r signif((acprop %>% filter(ACUsed==TRUE) %>% pull(p))*100,3)`% of U.S. households use A/C in their homes. If we wanted to know if this differs from 90%, we could set up our hypothesis as follows:

- $H_0: p = 0.90$ where $p$ is the proportion of the U.S. households that use A/C in their homes
- $H_A: p \neq 0.90$

To conduct this in R, we use the `svyttest()` function as follows:

```{r}
#| label: stattest-ttest-ex2
ttest_ex2 <- recs_des %>%
  svyttest(
    formula = (ACUsed == TRUE) - 0.90 ~ 0,
    design = .,
    na.rm = TRUE
  )

ttest_ex2
```

The output from the `svyttest()` function can be a bit hard to read. Using the {broom} package from tidymodels, a collection of packages for modeling using the tidyverse principles, we can clean up the output into a tibble to more easily understand what the test tells us.

```{r}
#| label: stattest-ttest-ex2-broom
broom::tidy(ttest_ex2)
```

The 'tidied' output can also be piped into the {gt} package to create a table ready for publication. We go over the {gt} package in Chapter \@ref(c08-communicating-results). The function `pretty_p_value()` comes from the {prettyunits} package and converts numeric p-values to characters and, by default prints four decimal places and displays any p-value less than 0.0001 as `"<0.0001"` though another minimum display p-value can be specified.

```{r}
#| label: stattest-ttest-ex2-gt
#| eval: FALSE
tidy(ttest_ex2) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

(ref:stattest-ttest-ex2-gt-tab)  One-sample t-test output for estimates of U.S. households use A/C in their homes differing from 90%, RECS 2020

```{r}
#| label: stattest-ttest-ex2-gt-tab
#| echo: FALSE
#| warning: FALSE

tidy(ttest_ex2) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number() %>%
  print_gt_book(knitr::opts_current$get()[["label"]])
```

The estimate differs from Example 1 in that the estimate is not displaying \(\mu - 0.90\) but rather \(\mu\), or the difference between the U.S. households that use A/C and the proportion we are comparing to. We can see that there is a difference of `r signif(ttest_ex2$estimate*100,3)` percentage points. Additionally, the t-statistic value in the `statistic` column is `r signif(ttest_ex2$statistic,3)`, and the p-value is `r pretty_p_value(ttest_ex2$p.value)`. These results indicate that the fewer than 90% of U.S. households use A/C in their homes.

#### Example 3: Unpaired two-sample t-test {.unnumbered #stattest-ttest-ex3}

Two additional variables in the RECS data are the electric bill cost (`DOLLAREL`) and whether the house used A/C or not (`ACUsed`).^[Is any air conditioning equipment used in your home?] If we want to know if the U.S. households that used A/C had higher electrical bills compared to those that did not, we could set up the hypothesis as follows:

- $H_0: \mu_{AC} = \mu_{noAC}$ where $\mu_{AC}$ is the electrical bill cost for U.S. households that used A/C and $\mu_{noAC}$ is the electrical bill cost for U.S. households that did not use A/C
- $H_A: \mu_{AC} \neq \mu_{noAC}$
 
Let's take a quick look at the data to see the format the data are in: 

```{r}
#| label: stattest-ttest-ex3-desc
recs_des %>%
  group_by(ACUsed) %>%
  summarize(mean = survey_mean(DOLLAREL, na.rm = TRUE))
```

To conduct this in R, we use `svyttest()`:

```{r stattest-ttest-ex3}
#| label: stattest-ttest-ex3
ttest_ex3 <- recs_des %>%
  svyttest(formula = DOLLAREL ~ ACUsed,
           design = .,
           na.rm = TRUE)

broom::tidy(ttest_ex3)
```

```{r}
#| label: stattest-ttest-ex3-gt
#| eval: FALSE
tidy(ttest_ex3) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

(ref:stattest-ttest-ex3-gt-tab)  Unpaired two-sample t-test output for estimates of U.S. households electrical bills by A/C use, RECS 2020

```{r}
#| label: stattest-ttest-ex3-gt-tab
#| echo: FALSE
#| warning: FALSE

tidy(ttest_ex3) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number() %>%
  print_gt_book(knitr::opts_current$get()[["label"]])
```

The results indicate that the difference in electrical bills for those that used A/C and those that did not is, on average, \$`r round(ttest_ex3$estimate,2)`. The difference appears to be statistically significant as the t-statistic is `r signif(ttest_ex3$statistic, 3)` and the p-value is $`r pretty_p_value(ttest_ex3[["p.value"]])`$. Households that used A/C spent, on average, $`r round(ttest_ex3[["estimate"]], 2) %>% unname()` more in 2020 on electricity than households without A/C.

#### Example 4: Paired two-sample t-test {.unnumbered #stattest-ttest-ex4}

Let's say we want to test whether the temperature that U.S. households set their thermostat at night differs depending on the season (comparing summer^[During the summer, what is your home’s typical indoor temperature inside your home at night?] and winter^[During the winter, what is your home’s typical indoor temperature inside your home at night?] temperatures). We could set up the hypothesis as follows:

- $H_0: \mu_{summer} = \mu_{winter}$ where $\mu_{summer}$ is the temperature that U.S. households set their thermostat to during summer nights, and $\mu_{winter}$ is the temperature that U.S. households set their thermostat to during winter nights
- $H_A: \mu_{summer} \neq \mu_{winter}$

To conduct this in R, we use `svyttest()` by calculating the temperature difference on the left-hand side as follows:

```{r}
#| label: stattest-ttest-ex4
ttest_ex4 <- recs_des %>%
  svyttest(
    design = .,
    formula = SummerTempNight - WinterTempNight ~ 0,
    na.rm = TRUE
  )

broom::tidy(ttest_ex4)
```

```{r}
#| label: stattest-ttest-ex4-gt
#| eval: FALSE
tidy(ttest_ex4) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number()
```

(ref:stattest-ttest-ex4-gt-tab)  Paired two-sample t-test output for estimates of U.S. households thermostat temperature by season, RECS 2020

```{r}
#| label: stattest-ttest-ex4-gt-tab
#| echo: FALSE
#| warning: FALSE

tidy(ttest_ex4) %>%
  mutate(p.value = pretty_p_value(p.value)) %>%
  gt() %>%
  fmt_number() %>%
  print_gt_book(knitr::opts_current$get()[["label"]])
```

U.S. households set their thermostat on average `r signif(ttest_ex4$estimate,2)`$^\circ$F warmer in summer nights than winter nights, which is statistically significant (t = `r signif(ttest_ex4$statistic, 3)`, p-value = $`r pretty_p_value(ttest_ex4[["p.value"]])`$).  

## Chi-Square Tests {#stattest-chi}

Chi-square tests ($\chi^2$) allow us to examine multiple proportions using a goodness-of-fit test, a test of independence, or a test of homogeneity. These three tests have the same $\chi^2$ distributions but with slightly different underlying assumptions.

First, **goodness-of-fit** tests are used when comparing *observed* data to *expected* data. For example, this could be used to determine if respondent demographics (the observed data in the sample) match known population information (the expected data). In this case, we can set up the hypothesis test as follows:

  - $H_0: p_1 = \pi_1, ~ p_2 = \pi_2, ~ ..., ~ p_k = \pi_k$ where $p_i$ is the observed proportion for category $i$, $\pi_i$ is expected proportion for category $i$, and $k$ is the number of categories
  - $H_A:$ at least one level of $p_i$ does not match $\pi_i$

Second, **tests of independence** are used when comparing two types of *observed* data to see if there is a relationship. For example, this could be used to determine if the proportion of respondents who voted for each political party in the presidential election matches the proportion of respondents who voted for each political party in a local election. In this case, we can set up the hypothesis test as follows:

  - $H_0:$ The two variables/factors are independent
  - $H_A:$ The two variables/factors are *not* independent 

Third, **tests of homogeneity** are used to compare two distributions to see if they match. For example, this could be used to determine if the highest education achieved is the same for both men and women. In this case, we can set up the hypothesis test as follows:

  - $H_0: p_{1a} = p_{1b}, ~ p_{2a} = p_{2b}, ~ ..., ~ p_{ka} = p_{kb}$ where $p_{ia}$ is the observed proportion of category $i$ for subgroup $a$, $p_{ib}$ is the observed proportion of category $i$ for subgroup $a$ and $k$ is the number of categories
  - $H_A:$ at least one category of $p_{ia}$ does not match $p_{ib}$

As with t-tests, the difference between using $\chi^2$ tests with non-survey data and survey data is based on the underlying variance estimation. The functions in the {survey} package account for these nuances, provided the design object is correctly defined. For basic variance estimation formulas for different survey design types, refer to Chapter \@ref(c10-specifying-sample-designs).

### Syntax {#stattest-chi-syntax}

When we do not have survey data, we may be able to use the `chisq.test()` function from the {stats} package. However, this function does not allow for weights or the variance structure to be accounted for with survey data. Therefore, when using survey data, we need to use one of two functions:

- `svygofchisq()`: For goodness of fit tests
- `svychisq()`: For tests of independence and homogeneity

The non-survey data function of `chisq.test()` requires either a single set of counts and given proportions (for goodness of fit tests) or two sets of counts for tests of independence and homogeneity. The functions we use with survey data require respondent-level data and formulas instead of counts. This ensures that the variances are correctly calculated.

First, the function for the goodness of fit tests is `svygofchisq()`:

```r
svygofchisq(formula,
            p,
            design,
            na.rm = TRUE,
            ...)
```

The arguments are:

* `formula`: Formula specifying a single factor variable
* `p`: Vector of probabilities for the categories of the factor in the correct order. If they probabilities do not sum to 1, they are rescaled to sum to 1.
* `design`: Survey design object
* ...: Other arguments to pass on, such as `na.rm`

Based on the order of the arguments, we again must use the dot `(.)` notation if we pipe in the survey design object or explicitly name the arguments as described in Section \@ref(dot-notation). For the goodness of fit tests, the formula is a single variable `formula = ~var` as we compare the observed data from this variable to the expected data. The expected probabilities are then entered in the `p` argument and need to be a vector of the same length as the number of categories in the variable. For example, if we want to know if the proportion of males and females matches a distribution of 30/70, then the sex variable (with two categories) would be used `formula = ~SEX`, and the proportions would be included as `p = c(.3, .7)`. It is important to note that the variable entered into the formula should be formatted as either a factor or a character. The examples below provide more detail and tips on how to make sure the levels match up correctly.

For tests of homogeneity and independence, the `svychisq()` function should be used. The syntax is as follows:

```r
svychisq(
  formula,
  design,
  statistic = c("F", "Chisq", "Wald", "adjWald",
                "lincom", "saddlepoint"),
  na.rm = TRUE
)
```

The arguments are:

* `formula`: Model formula specifying the table (shown in examples)
* `design`: Survey design object
* `statistic`: Type of test statistic to use in test (details below)
* `na.rm`: Remove missing values

There are six statistics that are accepted in this formula. For tests of homogeneity (when comparing cross-tabulations), the `F` or `Chisq` statistics should be used.^[These two statistics can also be used for goodness of fit tests if the `svygofchisq()` function is not used.] The `F` statistic is the default and uses the Rao-Scott second-order correction. This correction is designed to assist with complicated sampling designs (i.e., those other than a simple random sample) [@Scott2007]. The `Chisq` statistic is an adjusted version of the Pearson $\chi^2$ statistic. The version of this statistic in the `svychisq()` function compares the design effect estimate from the provided survey data to what the $\chi^2$ distribution would have been if the data came from a simple random sampling.

For tests of independence, the `Wald` and `adjWald` are recommended as they provide a better adjustment for variable comparisons [@lumley2010complex]. If the data have a small number of primary sampling units (PSUs) compared to the degrees of freedom, then the `adjWald` statistic should be used to account for this. The `lincom` and `saddlepoint` statistics are available for more complicated data structures.

The formula argument is always one-sided, unlike the `svyttest()` function. The two variables of interest should be included with a plus sign: `formula = ~ var_1 + var_2`. As with the `svygofchisq()` function, the variables entered into the formula should be formatted as either a factor or a character. 

Additionally, as with the t-test function, both `svygofchisq()` and `svychisq()` have the `na.rm` argument. If any data values are missing, the $\chi^2$ tests assume that `NA` is a category and include it in the calculation. Throughout this chapter, we always set `na.rm = TRUE`, but before analyzing the survey data, review the notes provided in Chapter \@ref(c11-missing-data) to better understand how to handle missing data.

### Examples {#stattest-chi-examples}

Let's walk through a few examples using the ANES data.

#### Example 1: Goodness of Fit Test {.unnumbered #stattest-chi-ex1}

ANES asked respondents about their highest education level.^[What is the highest level of school you have completed or the highest degree you have received?] Based on the data from the 2020 American Community Survey (ACS) 5-year estimates^[Data was pulled from data.census.gov using the S1501 Education Attainment 2020: ACS 5-Year Estimates Subject Tables], the education distribution of those aged 18+ in the United States (among the 50 states and District of Columbia) is as follows:

  - 11% had less than High School degree
  - 27% had a High School degree
  - 29% had some college or associate's degree
  - 33% had a bachelor's degree or higher

If we want to see if the weighted distribution from the ANES 2020 data matches this distribution, we could set up the hypothesis as follows:

  - $H_0: p_1 = 0.11, ~ p_2 = 0.27, ~ p_3 = 0.29, ~ p_4 = 0.33$
  - $H_A:$ at least one of the education levels does not match between the ANES and the ACS

To conduct this in R, let's first look at the education variable (`Education`) we have on the ANES data. Using the `survey_mean()` function discussed in Chapter \@ref(c05-descriptive-analysis), we can see the education levels and estimated proportions.

```{r}
#| label: stattest-chi-ex1-educmean
anes_des %>%
  drop_na(Education) %>%
  group_by(Education) %>%
  summarize(p = survey_mean())
```

Based on this output, we can see that we have different levels than the ACS data provide. Specifically, the education data from ANES include two levels for Bachelor's Degree or Higher (Bachelor's and Graduate), so these two categories need to be collapsed into a single category to match the ACS data. For this, among other methods, we can use the {forcats} package from the tidyverse. The package's `fct_collapse()` function helps us create a new variable by collapsing categories into a single one. Then, we use the `svygofchisq()` function to compare the ANES data to the ACS data where we specify the updated design object, the formula using the collapsed education variable, the ACS estimates for education levels as p, and removing NA values.

```{r}
#| label: stattest-chi-ex1
anes_des_educ <- anes_des %>%
  mutate(Education2 =
           fct_collapse(Education,
                        "Bachelor or Higher" = c("Bachelor's",
                                                 "Graduate")))

anes_des_educ %>%
  drop_na(Education2) %>%
  group_by(Education2) %>%
  summarize(p = survey_mean())

chi_ex1 <- anes_des_educ %>%
  svygofchisq(
    formula =  ~ Education2,
    p = c(0.11, 0.27, 0.29, 0.33),
    design = .,
    na.rm = TRUE
  )

chi_ex1
```

The output from the `svygofchisq()` indicates that at least one proportion from ANES does not match the ACS data ($\chi^2 =$ `r  prettyNum(chi_ex1$statistic, big.mark=",")`; p-value `r pretty_p_value(chi_ex1[["p.value"]])`). To get a better idea of the differences, we can use the `expected` output along with `survey_mean()` to create a comparison table:  

```{r}
#| label: stattest-chi-ex1-table
ex1_table <- anes_des_educ %>%
  drop_na(Education2) %>%
  group_by(Education2) %>%
  summarize(Observed = survey_mean(vartype = "ci")) %>%
  rename(Education = Education2) %>%
  mutate(Expected=c(0.11, 0.27, 0.29, 0.33)) %>%
  select(Education, Expected, everything())

ex1_table
```

This output includes our expected proportions from the ACS that we provided the `svygofchisq()` function along with the output of the observed proportions and their confidence intervals. This table shows that the "High school" and "Post HS" categories have nearly identical proportions but that the other two categories are slightly different. Looking at the confidence intervals, we can see that the ANES data skew to include fewer people in the "Less than HS" category and more people in the "Bachelor or Higher" category. This may be easier to see if we plot this. The code below uses the tabular output to create Figure \@ref(fig:stattest-chi-ex1-graph).

```{r}
#| label: stattest-chi-ex1-graph
#| fig.cap: Expected and observed proportions of education, showing the confidence intervals for the expected proportions and whether the observed proportions lie within them.
#| fig.alt: Expected and observed proportions of education, showing the confidence intervals for the expected proportions and whether the observed proportions lie within them. The x-axis has labels 'Less than HS', 'High school', 'Post HS', and 'Bachelor or Higher'. The only ones where expected proportion is outside of the intervals is 'Less than HS' and 'Bachelor or Higher'.

ex1_table %>%
  pivot_longer(
    cols = c("Expected", "Observed"),
    names_to = "Names",
    values_to = "Proportion"
  ) %>%
  mutate(
    Observed_low = if_else(Names == "Observed", Observed_low, NA_real_),
    Observed_upp = if_else(Names == "Observed", Observed_upp, NA_real_),
    Names = if_else(Names == "Observed", "ANES (observed)", "ACS (expected)")
  ) %>%
  ggplot(aes(x = Education, y = Proportion, color = Names)) +
  geom_point(alpha = 0.75, size = 2) +
  geom_errorbar(aes(ymin = Observed_low, ymax = Observed_upp), width = 0.25) +
  theme_bw() +
  scale_color_manual(name = "Type", values = book_colors[c(4, 1)]) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

#### Example 2: Test of Independence {.unnumbered #stattest-chi-ex2}

ANES asked respondents two questions about trust:

  - How often can you trust the federal government to do what is right?
  - How often can you trust other people?

If we want to see if the distributions of these two questions are similar or not, we can conduct a test of independence. Here is how the hypothesis could be set up:

  - $H_0:$ People's trust in the federal government and their trust in other people are independent (i.e., *not* related)
  - $H_A:$ People's trust in the federal government and their trust in other people are *not* independent (i.e., they are related)

To conduct this in R, we use the `svychisq()` function to compare the two variables:

```{r}
#| label: stattest-chi-ex2
chi_ex2 <- anes_des %>%
  svychisq(
    formula =  ~ TrustGovernment + TrustPeople,
    design = .,
    statistic = "Wald",
    na.rm = TRUE
  )

chi_ex2
```

The output from `svychisq()` indicates that the distribution of people's trust in the federal government and their trust in other people are *not* independent, meaning that they are related. Let's output the distributions in a table to see the relationship. The `observed` output from the test provides a cross-tabulation of the counts for each category:

```{r}
#| label: stattest-chi-ex2-counts
chi_ex2$observed
```

However, as researchers, we often want to know about the proportions and not just the respondent counts from the survey. There are a couple of different ways that we can do this. The first is using the counts from `chi_ex2$observed` to calculate the proportion. We can then pivot the table to create a cross-tabulation similar to the counts table above. Adding `group_by()` to the code means that we are obtaining the proportions within each level of that variable. In this case, we are looking at the distribution of `TrustGovernment` for each level of `TrustPeople`. The resulting table is shown in Table \@ref(tab:stattest-chi-ex2-prop1-tab).

```{r}
#| label: stattest-chi-ex2-prop1
#| warning: false
chi_ex2_table<-chi_ex2$observed %>% 
  as_tibble() %>%
  group_by(TrustPeople) %>%
  mutate(prop = round(n / sum(n), 3)) %>%
  select(-n) %>%
  pivot_wider(names_from = TrustPeople, values_from = prop) %>% 
  gt(rowname_col = "TrustGovernment") %>%
  tab_stubhead(label = "Trust in Government") %>%
  tab_spanner(label = "Trust in People",
              columns = everything()) %>%
  cols_label(`Most of the time` = md("Most of<br />the time"),
             `About half the time` = md("About half<br />the time"),
             `Some of the time` = md("Some of<br />the time"))
```

```{r}
#| label: stattest-chi-ex2-prop1-noeval
#| eval: false
chi_ex2_table
```

(ref:stattest-chi-ex2-prop1-tab) Proportion of adults in the U.S. by levels of trust in people and government, ANES 2020

```{r}
#| label: stattest-chi-ex2-prop1-tab
#| echo: FALSE
#| warning: FALSE

chi_ex2_table %>%
    print_gt_book(knitr::opts_current$get()[["label"]])
```

In Table \@ref(tab:stattest-chi-ex2-prop1-tab), each column sums to 1. For example, we can say that it is estimated that of people who always trust in people, `r round(chi_ex2$observed[1,1]/sum(chi_ex2$observed[,1])*100, 1)`% also always trust in government based on the top-left cell but `r round(chi_ex2$observed[5,1]/sum(chi_ex2$observed[,1])*100, 1)`% never trust in government.

The second option is to use `group_by()` and `survey_mean()` functions to calculate the proportions from the ANES design object. A reminder that with more than one variable listed in the `group_by()` statement, the proportions are within the first variable listed. As mentioned above, we are looking at the distribution of `TrustGovernment` for each level of `TrustPeople`. 
```{r}
#| label: stattest-chi-ex2-prop2
chi_ex2_obs <- anes_des %>%
  drop_na(TrustPeople, TrustGovernment) %>%
  group_by(TrustPeople, TrustGovernment) %>%
  summarize(Observed = round(survey_mean(vartype = "ci"), 3),
            .groups="drop") 

chi_ex2_obs_table<-chi_ex2_obs %>%
  mutate(prop = paste0(Observed, " (", Observed_low, ", ",
                       Observed_upp, ")")) %>%
  select(TrustGovernment, TrustPeople, prop) %>%
  pivot_wider(names_from = TrustPeople, values_from = prop) %>%
  gt(rowname_col = "TrustGovernment") %>%
  tab_stubhead(label = "Trust in Government") %>%
  tab_spanner(label = "Trust in People",
              columns = everything()) %>% 
  tab_options(page.orientation = "landscape")
```

```{r}
#| label: stattest-chi-ex2-prop2-noeval
#| eval: false
chi_ex2_obs_table
```

(ref:stattest-chi-ex2-prop2-tab) Proportion of adults in the U.S. by levels of trust in people and government with confidence intervals, ANES 2020

```{r}
#| label: stattest-chi-ex2-prop2-tab
#| echo: FALSE
#| warning: FALSE

chi_ex2_obs_table %>%
    print_gt_book(knitr::opts_current$get()[["label"]])
```

Both methods produce the same output as the `svychisq()` function does account for the survey design. However, calculating the proportions directly from the design object means we can also obtain the variance information. In this case, the table output displays the survey estimate followed by the confidence intervals. Based on the output, we can see that of those who never trust people, `r round(chi_ex2$observed[5,5]/sum(chi_ex2$observed[,5])*100, 1)`% also never trust the government, while the proportions of never trusting the government are much lower for each of the other levels of trusting people. 

We may find it easier to look at these proportions graphically. We can use `ggplot()` and facets to provide an overview as shown below to create Figure \@ref(fig:stattest-chi-ex2-graph):

<!--Perhaps a mosaic plot instead?-->
```{r}
#| label: stattest-chi-ex2-graph
#| fig.cap: Proportion of adults in the U.S. by levels of trust in people and government with confidence intervals, ANES 2020
#| fig.alt: Proportion of adults in the U.S. by levels of trust in people and government with confidence intervals, ANES 2020. This presents the same information as the previous table in graphical form.

chi_ex2_obs %>%
  mutate(TrustPeople=
           fct_reorder(str_c("Trust in People:\n", TrustPeople), 
                       order(TrustPeople))) %>%
  ggplot(aes(x = TrustGovernment, y = Observed, color = TrustGovernment)) +
  facet_wrap( ~ TrustPeople, ncol = 5) +
  geom_point() +
  geom_errorbar(aes(ymin = Observed_low, ymax = Observed_upp)) +
  ylab("Proportion") +
  xlab("") +
  theme_bw() +
  scale_color_manual(name="Trust in Government", values=book_colors) +
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  guides(col = guide_legend(nrow=2))
```

#### Example 3: Test of Homogeneity {.unnumbered #stattest-chi-ex3}

Researchers and politicians often look at specific demographics each election cycle to understand how each group is leaning or voting toward candidates. The ANES data are collected post-election, but we can still see if there are differences in how specific demographic groups voted.

If we want to see if there is a difference in how each age group voted for the 2020 candidates, this would be a test of homogeneity, and we can set up the hypothesis as follows:

\begin{align*}
H_0: p_{1_{Biden}} &= p_{1_{Trump}} = p_{1_{Other}},\\
    p_{2_{Biden}} &= p_{2_{Trump}} = p_{2_{Other}},\\
    p_{3_{Biden}} &= p_{3_{Trump}} = p_{3_{Other}},\\
    p_{4_{Biden}} &= p_{4_{Trump}} = p_{4_{Other}},\\
    p_{5_{Biden}} &= p_{5_{Trump}} = p_{5_{Other}},\\
    p_{6_{Biden}} &= p_{6_{Trump}} = p_{6_{Other}}
  \end{align*}
    where $p_{i_{Biden}}$ is the observed proportion of each age group ($i$) that voted for Joseph Biden, $p_{i_{Trump}}$ is the observed proportion of each age group ($i$) that voted for Donald Trump, and $p_{i_{Other}}$ is the observed proportion of each age group ($i$) that voted for another candidate
    
- $H_A:$ at least one category of $p_{i_{Biden}}$ does not match $p_{i_{Trump}}$ or $p_{i_{Other}}$

To conduct this in R, we use the `svychisq()` function to compare the two variables:

```{r}
#| label: stattest-chi-ex3
chi_ex3 <- anes_des %>%
  drop_na(VotedPres2020_selection, AgeGroup) %>%
  svychisq(
    formula =  ~ AgeGroup + VotedPres2020_selection,
    design = .,
    statistic = "Chisq",
    na.rm = TRUE
  )

chi_ex3
```

The output from `svychisq()` indicates a difference in how each age group voted in the 2020 election. To get a better idea of the different distributions, let's output proportions to see the relationship. As we learned in Example 2 above, we can use `chi_ex3$observed`, or if we want to get the variance information (which is crucial with survey data), we can use `survey_mean()`. Remember, when we have two variables in `group_by()`, we obtain the proportions within each level of the variable listed. In this case, we are looking at the distribution of `AgeGroup` for each level of `VotedPres2020_selection`.

```{r}
#| label: stattest-chi-ex3-table
chi_ex3_obs <- anes_des %>%
  filter(VotedPres2020 == "Yes") %>%
  drop_na(VotedPres2020_selection, AgeGroup) %>%
  group_by(VotedPres2020_selection, AgeGroup) %>%
  summarize(Observed = round(survey_mean(vartype = "ci"), 3)) 

chi_ex3_obs_table<-chi_ex3_obs %>%
  mutate(prop = paste0(Observed, " (", Observed_low, ", ",
                       Observed_upp, ")")) %>%
  select(AgeGroup, VotedPres2020_selection, prop) %>%
  pivot_wider(names_from = VotedPres2020_selection, 
              values_from = prop) %>%
  gt(rowname_col = "AgeGroup") %>%
  tab_stubhead(label = "Age Group")
```

```{r}
#| label: stattest-chi-ex3-table-noeval
#| eval: false
chi_ex3_obs_table
```

(ref:stattest-chi-ex3-tab) Distribution of age group by presidential candidate selection with confidence intervals

```{r}
#| label: stattest-chi-ex3-tab
#| echo: FALSE
#| warning: FALSE

chi_ex3_obs_table %>%
    print_gt_book(knitr::opts_current$get()[["label"]])
```

We can see that the age group distribution that voted for Biden and other candidates was younger than those that voted for Trump. For example, of those who voted for Biden, 20.4% were in the 18-29 age group, compared to only 11.4% of those who voted for Trump were in that age group. On the other side, 23.4% of those who voted for Trump were in the 50-59 age group compared to only 15.4% of those who voted for Biden.

## Exercises {#stattest-exercises}

The exercises use the design objects `anes_des` and `recs_des` as provided in the Prerequisites box in the [beginning of the chapter](#c06-statistical-testing). Here are some exercises for practicing conducting t-tests using `svyttest()`:

1. Using the RECS data, do more than 50% of U.S. households use A/C (`ACUsed`)?

2. Using the RECS data, does the average temperature that U.S. households set their thermostats to differ between the day and night in the winter (`WinterTempDay` and `WinterTempNight`)?

3. Using the ANES data, does the average age (`Age`) of those who voted for Joseph Biden in 2020 (`VotedPres2020_selection`) differ from those who voted for another candidate?

4. If we wanted to determine if the political party affiliation differed for males and females, what test would we use?

  a. Goodness of fit test (`svygofchisq()`)
  b. Test of independence (`svychisq()`)
  c. Test of homogeneity (`svychisq()`)

5. In the RECS data, is there a relationship between the type of housing unit (`HousingUnitType`) and the year the house was built (`YearMade`)?

6. In the ANES data, is there a difference in the distribution of gender (`Gender`) across early voting status in 2020 (`EarlyVote2020`)?