# (PART) Reporting {-}

# Communicating results {#c08-communicating-results}

```{r}
#| include: false
knitr::opts_chunk$set(tidy = 'styler')
```

::: {.prereqbox-header}
`r if (knitr:::is_html_output()) '### Prerequisites {- #prereq8}'`
:::

::: {.prereqbox data-latex="{Prerequisites}"}
For this chapter, load the following packages:
```{r}
#| label: results-setup
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(survey) 
library(srvyr) 
library(srvyr.data)
library(gt)
library(gtsummary)
```

We will be using data from ANES as described in Chapter \@ref(c04-set-up). As a reminder, here is the code to create the design objects for each to use throughout this chapter. For ANES, we need to adjust the weight so it sums to the population instead of the sample (see the ANES documentation and Chapter \@ref(c04-set-up) for more information).

```{r}
#| label: results-anes-des
#| eval: FALSE
targetpop <- 231592693
data(anes_2020)

anes_adjwgt <- anes_2020 %>%
  mutate(Weight = Weight / sum(Weight) * targetpop)

anes_des <- anes_adjwgt %>%
  as_survey_design(
    weights = Weight,
    strata = Stratum,
    ids = VarUnit,
    nest = TRUE
  )
```
:::

## Introduction

After finishing the analysis and modeling, we then proceed to the important task of communicating the survey results. Our audience may range from seasoned researchers familiar with our survey data to newcomers encountering the information for the first time. We should aim to explain the methodology and analysis while presenting findings in an accessible way, and it is our responsibility to report information with care.

Before beginning any dissemination of results, consider questions such as:

  - How will we present results? Examples include a website, print, or other media. Based on the media type, we might limit or enhance the use of graphical representation. 
  - What is the audience's familiarity with the study and/or data? Audiences can range from the general public to data experts. If we anticipate limited knowledge about the study, we should provide detailed descriptions (we discuss recommendations later in the chapter).
  - What are we trying to communicate? It could be summary statistics, trends, patterns, or other insights. Tables might suit summary statistics, while plots are better at conveying trends and patterns.
  - Is the audience accustomed to interpreting plots? If not, include explanatory text to guide them on how to interpret the plots effectively.
  - What is the audience's statistical knowledge? If the audience does not have a strong statistics background, provide text on standard errors, confidence intervals, and other estimate types to enhance understanding.

## Describing results through text

As analysts, our emphasis is often on the data, and communicating results can sometimes be overlooked. However, we need to identify the appropriate information to share with our audience. Chapters \@ref(c02-overview-surveys) and \@ref(c03-understanding-survey-data-documentation) provide insights into factors we need to consider during analysis, and they remain relevant when presenting results to others.

### Methodology

If we are using existing data, methodologically-sound surveys will provide documentation about how the survey was fielded, the questionnaires, and other necessary information for analyses. For example, the survey's methodology reports should include the population of interest, sampling procedures, response rates, questionnaire documentation, weighting, and a general overview of disclosure statements. Many American organizations follow the American Association for Public Opinion Research's (AAPOR) [Transparency Initiative](https://aapor.org/standards-and-ethics/transparency-initiative). The AAPOR Transparency Initiative require organizations to include specific details in their methodology, making it clear how we can and should analyze the results. Being transparent about these methods is vital for the scientific rigor of the field.

The details provided in Chapter \@ref(c02-overview-surveys) about for researchers should be shared with the audience when presenting the results. When using publicly-available data, like the examples in this book, we can often link to the methodology report in our final output. We should also provide high-level information for the audience to quickly grasp the context around the findings. For example, we can mention when and where the study was conducted, the population's age range, or other contextual details. This information helps the audience understand how generalizable the results are.

Providing this material is especially important when there's no methodology report available for the analyzed data. For example, if a researcher conducted a new survey for a specific purpose, we should document and present all the pertinent information during the analysis and reporting process. Adhering to the AAPOR Transparency Initiative guidelines is a reliable method to guarantee that all essential information is communicated to the audience.

### Analysis

Along with the survey methodology and weight calculations, we should also share our approach to preparing, cleaning, and analyzing the data. For example, in Chapter \@ref(c06-statistical-testing), we compared education distributions from the ANES survey to the American Community Survey (ACS). To make the comparison, we had to collapse education categories provided in the ANES data to match the ACS. The process for this particular example may seem straightforward (like combining Bachelor's and Graduate Degrees into a single category), but there are multiple ways to deal with the data. Our choice is just one of many. We should document both the original ANES question and response options and the steps we took to match it with ACS data. This transparency helps clarify our analysis to our audience.

Missing data is another instance where we want to be unambigious and upfront with our audience. In this book, numerous examples and exercises remove missing data, as this is often the easiest way to handle them. However, there are circumstances where missing data holds substantive importance, and excluding them could introduce bias (see Chapter \@ref(c11-missing-data)). Being transparent about our handling of missing data is important to maintaining the integrity of our analysis and ensuring a comprehensive understanding of the results.

### Results

While tables and graphs are commonly used to communicate results, there are instances where text can be more effective in sharing information. Narrative details, such as context around point estimates or model coefficients, can go a long way in improving our communication. We have several strategies to effectively convey the significance of the data to the audience through text.

First, we can highlight important data points in a sentence using plain language. For example, if we were looking at election polling data conducted before an election, we could say something like:  

> As of [DATE], an estimated XX% of registered U.S. voters say they will vote for [CANDIDATE NAME] for president in the [YEAR] general election.

This sentence provides key pieces of information in a straightforward way:

 1. **[DATE]**: Given that polling data is time-specific, providing the date of reference lets the audience know when this data was valid.
 2. **Registered U.S. voters**: This tells the audience who we surveyed, letting them know the target population.
 3. **XX%**: This part provides the estimated percentage of people voting for a specific candidate for a specific office.
 4. **[YEAR] general election**: As with the bullet above, adding this gives more context about the election type and year. The estimate would take on a different meaning if we changed it to a *primary* election instead of a *general* election.
 
We also included the word "estimated." When presenting aggregate survey results, we have errors around each estimate. We want to convey this uncertainty rather than talk in absolutes. Words like "estimated," "on average," or "around" can help our communication to the audience. Instead of saying 'XX%,' we can say 'XX% (+/- Y%)' to show the margin of error. Confidence intervals can also be incorporated into the text to assist readers.  

Second, providing context and discussing the *meaning* behind a point can help the audience glean some insight into why the data is important. For example, when comparing two points, it can be helpful to highlight if there are statistically significant differences and explain the impact and relevance of this information. This is where we, as analysts, should to do our best to be mindful of biases and present the facts logically. 

Keep in mind how we discuss these findings can greatly influence how the audience interprets them. If we include speculation, using phrases like "the authors speculate" or "these findings may indicate" relays the uncertainty around the notion while still lending a plausible solution. Additionally, we can present alternative viewpoints or competing discussion points to explain the uncertainty in the results.

## Visualizing data

Although discussing key findings in the text is important, presenting large amounts of data is often more digestible for the audience in tables or visualizations. Effectively combining text, tables, and graphs can be powerful in communicating results. This section provides examples of using the {gt}, {gtsummary}, and {ggplot2} packages to enhance the dissemination of results.

### Tables

Tables are a great way to provide a large amount of data when individual data points need to be examined. However, it is important to present tables in a reader-friendly format. Numbers should align, rows and columns should be easy to follow, and the table size should not compromise readability. Using key visualization techniques, we can create tables that are informative and nice to look at. Many packages create easy-to-read tables (e.g., {kable} \+ {kableExtra}, {gt}, {gtsummary}, {DT}, {formattable}, {flextable}, {reactable}). While we fill focus on {gt} here, we encourage learning about others as they may have additional helpful features. We appreciate the flexibility, pipability, and numerous extensions of the {gt} package. Please note, at this time, {gtsummary} needs additional features to be widely used for survey analysis, particularly its lack of ability to work with replicate designs. We provide one example using {gtsummary} and hope it evolves into a more comprehensive tool over time.

#### Transitioning {srvyr} output to a {gt} table {#results-gt}

Let's start by using some of the data we calculated earlier in this book. In Chapter \@ref(c06-statistical-testing), we looked at data on trust in government with the proportions calculated below:

```{r}
#| label: results-table-raw
trust_gov <- anes_des %>%
  drop_na(TrustGovernment) %>%
  group_by(TrustGovernment) %>%
  summarize(trust_gov_p = survey_prop())

trust_gov
```

The default output generated by R may work for initial viewing inside RStudio or when creating basic output in an R Markdown or Quarto document. However, when presenting these results in other publications, such as the print version of this book or for more formal dissemination, modifying the display can improve our reader's experience. 

Looking at the output from `trust_gov`, there are a couple of items that are probably obvious to fix: (1) use percentages instead of proportions and (2) the variable names as the column headers. The {gt} package is a good tool for implementing better labeling and creating publishable tables. Let's walk through some code as we implement a few changes to improve the table's usefulness. 

We begin with the `gt()` function to initiate the table and use the argument `rowname_col()` to make the `TrustGovernment` column the labels for each row (called the table "stub"). The `cols_label()` function is used to create informative column labels instead of variable names, and the `tab_spanner()` function is applied to add a label across multiple columns. In this case, we apply the label "Trust in Government, 2020" across all the columns except the stub. Finally, the `fmt_percent()` function is used to format the proportions into percentages and reduce the number of decimals shown. Note, the `tab_caption()` function is used to add a table title for the book in HTML and allows for cross-referencing in R Markdown Quarto and bookdown, as well as adding it to the list of tables in the book.

```{r}
#| label: results-table-gt1
trust_gov_gt <- trust_gov %>%
  gt(rowname_col = "TrustGovernment") %>%
  cols_label(trust_gov_p = "%",
             trust_gov_p_se = "s.e. (%)") %>%
  tab_spanner(label = "Trust in Government, 2020",
              columns = c(trust_gov_p, trust_gov_p_se)) %>%
  fmt_percent(decimals = 1)
```

```{r}
#| label: results-table-gt1-noeval
#| eval: false
trust_gov_gt %>% 
  tab_caption("Example of gt table with trust in government estimate")
```

(ref:results-table-gt1-tab) Example of gt table with trust in government estimate

```{r}
#| label: results-table-gt1-out
#| echo: FALSE
#| warning: FALSE

trust_gov_gt %>%
  print_gt_book("results-table-gt1-tab")
```

A few more things we can add are a title, a data source note, and a footnote with the question information using the functions `tab_header()`, `tab_source_note()`, and `tab_footnote()`. We can decide that having the percentage sign in both the header and the cells is redundant, so instead of using `fmt_percent()`, we use `fmt_number()` and scale the number by 100.

```{r}
#| label: results-table-gt2
trust_gov_gt2 <- trust_gov_gt %>%
  tab_header("American voter's trust
             in the federal government, 2020") %>%
  tab_source_note("American National Election Studies, 2020") %>%
  tab_footnote(
    "Question text: How often can you trust the federal government
    in Washington to do what is right?"
  ) %>%
  fmt_number(scale_by = 100,
             decimals = 1)
```

```{r}
#| label: results-table-gt2-noeval
#| eval: false
trust_gov_gt2
```

(ref:results-table-gt2-tab) Example of gt table with trust in government estimates with additional context

```{r}
#| label: results-table-gt2-out
#| echo: FALSE
#| warning: FALSE

trust_gov_gt2 %>%
  print_gt_book("results-table-gt2-tab")
```

#### Expanding Tables using {gtsummary} {-}

The {gtsummary} package simultaneously summarizes data and creates publication-ready tables. Its origins are in clinical trial data but it has been extended to include survey analysis in some limited ways. At this time, it only works with survey objects using Taylor's Series Linearization and not replicate methods. A limited set of summary statistics are available. For categorical variables, the following summary statistics are available:

  - `{n}` frequency
  - `{N}` denominator, or cohort size
  - `{p}` percentage
  - `{p.std.error}` standard error of the sample proportion
  - `{deff}` design effect of the sample proportion
  - `{n_unweighted}` unweighted frequency
  - `{N_unweighted}` unweighted denominator
  - `{p_unweighted}` unweighted formatted percentage

For continuous variables, the following summary statistics are available:

  - `{median}` median
  - `{mean}` mean
  - `{mean.std.error}` standard error of the sample mean
  - `{deff}` design effect of the sample mean
  - `{sd}` standard deviation
  - `{var}` variance
  - `{min}` minimum
  - `{max}` maximum
  - `{p##}` any integer percentile, where `##` is an integer from 0 to 100
  - `{sum}` sum

In the following example, we will build up a table using {gtsummary}, which will be similar to the table in the {gt} example. The main function used is `tbl_svysummary()`. In this function, the variables we want to analyze are included in the `include` argument, and the statistics we want to display are in the `statistic` argument. To specify statistics, the syntax from the {glue} package is used where variables you want to insert are included inside curly brackets. To specify that we want, the proportion followed by the standard error of the proportion in parentheses, we use "{p} ({p.std.error})". We must specify the statistics we want using the names of the statistics in the two lists above.

```{r}
#| label: results-gts-ex-1
anes_des_gtsum<-anes_des %>%
  tbl_svysummary(include = TrustGovernment,
                 statistic = list(all_categorical() ~ "{p} ({p.std.error})")) 
```

```{r}
#| label: results-table-gt3-noeval
#| eval: false
anes_des_gtsum
```

(ref:results-gts-ex-1-tab) Example of gtsummary table with trust in government estimates

```{r}
#| label: results-gts-ex-1-out
#| echo: FALSE
#| warning: FALSE

anes_des_gtsum %>% 
  print_gt_book("results-gts-ex-1-tab")
```

In this default table, the weighted number of missing (or Unknown) records is included. Additionally, the standard error is reported as a proportion while the proportion is styled as a percentage. In the next step, we remove the Unknown category by setting the missing argument to "no" and format the standard error as a percentage within the digits argument. Finally, we label the "TrustGovernment" variable to something more publication-ready using the label argument.

```{r}
#| label: results-gts-ex-2
anes_des_gtsum2<-anes_des %>%
  tbl_svysummary(
    include = TrustGovernment,
    statistic = list(all_categorical() ~ "{p} ({p.std.error})"),
    missing = "no",
    digits = list(TrustGovernment ~ style_percent),
    label = list(TrustGovernment ~ "Trust in Government, 2020")
  )
```

```{r}
#| label: results-gts-ex-2-noeval
#| eval: false
anes_des_gtsum2
```

(ref:results-gts-ex-2-tab) Example of gtsummary table with trust in government estimates with labeling and digits options

```{r}
#| label: results-gts-ex-2-out
#| echo: FALSE
#| warning: FALSE

anes_des_gtsum2 %>%
  print_gt_book("results-gts-ex-2-tab")
```

To remove the phrase "Characteristic" and the estimated population size, we can modify the header using the function `modify_header()` to update the label and stat_0. To add footnotes and a title, we do this after converting the object to a gt table using `as_gt()` and can use the same functions we did in Section \@ref(results-gt).

```{r}
#| label: results-gts-ex-3
anes_des_gtsum3<-anes_des %>%
  tbl_svysummary(
    include = TrustGovernment,
    statistic = list(all_categorical() ~ "{p} ({p.std.error})"),
    missing = "no",
    digits = list(TrustGovernment ~ style_percent),
    label = list(TrustGovernment ~ "Trust in Government, 2020")
  ) %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_header(label = " ",
                stat_0 = "% (s.e.)") %>%
  as_gt() %>%
  tab_header("American voter's trust 
             in the federal government, 2020") %>%
  tab_source_note("American National Election Studies, 2020") %>%
  tab_footnote(
    "Question text: How often can you trust 
    the federal government in Washington 
    to do what is right?"
  )
```

```{r}
#| label: results-gts-ex-3-noeval
#| eval: false
anes_des_gtsum3
```

(ref:results-gts-ex-3-tab) Example of gtsummary table with trust in government estimates with more labeling options and context

```{r}
#| label: results-gts-ex-3-out
#| echo: FALSE
#| warning: FALSE

anes_des_gtsum3 %>%
  print_gt_book("results-gts-ex-3-tab")
```

Continuous variables can also be added, and we add a summary of the age variable to the table below by updating the include, statistic, and digits argument. Adding on additional variables is a large benefit to the {gtsummary} package.

```{r}
#| label: results-gts-ex-4
#| tidy: FALSE
anes_des_gtsum4<-anes_des %>%
  tbl_svysummary(
    include = c(TrustGovernment, Age),
    statistic = list(
      all_categorical() ~ "{p} ({p.std.error})",
      all_continuous() ~ "{mean} ({mean.std.error})"
    ),
    missing = "no",
    digits = list(TrustGovernment ~ style_percent,
                  Age ~ c(1, 2)),
    label = list(TrustGovernment ~ "Trust in Government, 2020")
  ) %>%
  modify_header(label = " ",
                stat_0 = "Summary") %>%
  as_gt() %>%
  tab_header("American voter's trust in the federal government, 2020") %>%
  tab_source_note("American National Election Studies, 2020") %>%
  tab_footnote(
    "Question text: How often can you trust the federal government
    in Washington to do what is right?"
  ) %>%
  tab_caption("Example of gtsummary table with trust in government 
              estimates and average age")
```

```{r}
#| label: results-gts-ex-4-noeval
#| eval: false
anes_des_gtsum4
```

(ref:results-gts-ex-4-tab) Example of gtsummary table with trust in government estimates and average age

```{r}
#| label: results-gts-ex-4-out
#| echo: FALSE
#| warning: FALSE

anes_des_gtsum4 %>%
  print_gt_book("results-gts-ex-4-tab")
```

The {gtsummary} also allows calculating statistics by different groups  easily. Let's adapt the prior example to perform analysis by whether the person voted for president in 2020. The argument for by is updated and the header names are updated. Finally, we update the header.

```{r}
#| label: results-gts-ex-5
#| messages: FALSE
anes_des_gtsum5<-anes_des %>%
  drop_na(VotedPres2020) %>%
  tbl_svysummary(
    include=TrustGovernment,
    statistic = list(all_categorical() ~ "{p} ({p.std.error})"),
    missing="no",
    digits = list(TrustGovernment ~ style_percent),
    label=list(TrustGovernment ~ "Trust in Government, 2020"),
    by = VotedPres2020
  ) %>%
  modify_header(
    label = " ",
    stat_1 = "Voted",
    stat_2 = "Didn't vote"
  ) %>%
  as_gt() %>%
  tab_header("American voter's trust 
             in the federal government by whether they voted 
             in the 2020 presidential election") %>%
  tab_source_note("American National Election Studies, 2020") %>%
  tab_footnote("Question text: How often can you trust the federal government 
               in Washington to do what is right?")
```

```{r}
#| label: results-gts-ex-5-noeval
#| eval: false
anes_des_gtsum5
```

(ref:results-gts-ex-5-tab) Example of gtsummary table with trust in government estimates by voting status

```{r}
#| label: results-gts-ex-5-out
#| echo: FALSE
#| warning: FALSE

anes_des_gtsum5 %>%
    print_gt_book("results-gts-ex-5-tab")
```


### Charts and Plots

Survey analysis can result in an abundance of printed summary statistics and models. Even with the best analysis, the results can be overwhelming and difficult to comprehend. This is where charts and plots play a key role in our work. By transforming complex data into a visual representation, we can recognize patterns, relationships, and trends with greater ease.

R has many packages for creating compelling and insightful charts. We will focus on {ggplot2}, a member of the {tidyverse}. This package is a powerful, flexible tool for creating a wide range of data visualization.

{ggplot2} follows the "grammar of graphics," a framework that incrementally adds layers of chart components. We can customize visual elements such as scales, colors, labels, and annotations to enhance the understanding of data. After creating the design object that we've been using previously, we select our desired data points by modifying the existing design to add other outcomes and calculate estimates. Below, we create a binary variable `TrustGovernmentUsually`, which is TRUE when TrustGovernment is "Always" or "Most of the time" and FALSE otherwise. Then, we calculate the percentage of people who usually trust the government by who they voted for in the 2020 presidential election (`VotedPres2020_selection`). We remove the cases where people did not vote or did not indicate for whom they voted.

```{r}
#| label: results-anes-prep
anes_des_der <- anes_des %>%
  mutate(TrustGovernmentUsually = case_when(
    is.na(TrustGovernment) ~ NA,
    TRUE ~ TrustGovernment %in% c("Always", "Most of the time")
  )) %>%
  drop_na(VotedPres2020_selection) %>%
  group_by(VotedPres2020_selection) %>%
  summarize(
    pct_trust = survey_mean(
      TrustGovernmentUsually,
      na.rm = TRUE,
      proportion = TRUE,
      vartype = "ci"
    ),
    .groups = "drop"
  )

anes_des_der
```

Now, we can begin creating our chart with {ggplot2}. First, we set up our plot with `ggplot()`. Next, we state the data points we want to show with aesthetics, or `aes`.  Aesthetics are visual properties of the objects in your plot. In the example below, we map the `x` variable to `VotedPres2020_selection` from the dataset, and the `y` variable to `pct_trust`. Finally, we specify the type of plot with `geom_*()`, in this case, `geom_bar()`. This initial plot is displayed in Figure \@ref(fig:results-plot1).

```{r}
#| label: results-plot1
#| fig.cap: "Bar chart of trust in government by chosen 2020 presidential candidate"
#| fig.alt: "Bar chart with x-axis of 'VotedPres2020_selection' with labels Biden, Trump and Other. It has y-axis 'pct_trust' with labels 0.00, 0.05, 0.10 and 0.15. The chart is a bar chart with 3 vertical bars. Bar 1 (Biden) has a height of 0.12. Bar 2 (Trump) has a height of 0.17. Bar 3 (Other) has a height of 0.06."
p <- anes_des_der %>%
  ggplot(aes(x = VotedPres2020_selection,
             y = pct_trust)) +
  geom_bar(stat = "identity")

p
```

This is a great starting point: we can see that the percentage of people saying they usually trust the government is higher for those who voted for Trump than Biden or other candidates. What if we wanted to add color to better differentiate the three groups? We can add `fill` under `aesthetics`, denoting that we want to use the different values of `VotedPres2020_selection` to fill in the bars. In this example, different colors will be used for Biden and Trump.

```{r}
#| label: results-plot2
#| fig.cap: "Bar chart of trust in government by chosen 2020 presidential candidate with colors"
#| fig.alt: "Bar chart with x-axis of 'VotedPres2020_selection' with labels Biden, Trump and Other. It has y-axis 'pct_trust' with labels 0.00, 0.05, 0.10 and 0.15. The chart is a bar chart with 3 vertical bars. Bar 1 (Biden) has a height of 0.12 and a color of strong reddish orange. Bar 2 (Trump) has a height of 0.17 and a color of vivid yellowish green. Bar 3 (Other) has a height of 0.06 and color of brilliant blue."
pcolor <- anes_des_der %>%
  ggplot(aes(x = VotedPres2020_selection,
             y = pct_trust,
             fill = VotedPres2020_selection)) +
  geom_bar(stat = "identity")

pcolor
```

Let's say we wanted to follow good statistical analysis practice and include our variability in our plot. We can add another geom, `geom_errorbar()`, to display the confidence intervals on top of our `geom_bar()` layer using a plus sign `+`. 

```{r}
#| label: results-plot3
#| fig.cap: "Bar chart of trust in government by chosen 2020 presidential candidate with colors and error bars"
#| fig.alt: "Bar chart with x-axis of 'VotedPres2020_selection' with labels Biden, Trump and Other. It has y-axis 'pct_trust' with labels 0.00, 0.05, 0.10 and 0.15. The chart is a bar chart with 3 vertical bars. Bar 1 (Biden) has a height of 0.12 and a color of strong reddish orange. Bar 2 (Trump) has a height of 0.17 and a color of vivid yellowish green. Bar 3 (Other) has a height of 0.06 and color of brilliant blue. Error bars are added with the Bar 1 (Biden) error ranging from 0.11 to 0.14, Bar 2 (Trump) error ranging from 0.16 to 0.19, and the Bar 3 (Other) error ranging from 0.02 to 0.14."
pcol_error <- anes_des_der %>%
  ggplot(aes(x = VotedPres2020_selection,
             y = pct_trust,
             fill = VotedPres2020_selection)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = pct_trust_low,
                    ymax = pct_trust_upp),
                width = .2)

pcol_error
```

We can continue adding to our plot until we achieve the visualization we'd like to present. For example, we can remove the color legend since it does not add any useful information with `guides(fill = "none")`. We can specify values for `fill` using `scale_fill_manual()`. Inside the function, we provide a vector of values for each of the colors in our plot. These values are hexadecimal (hex) color codes, denoted by a leading pound sign `#` followed by six letters or numbers. The hex code `#0b3954` is a dark blue. There are many tools online that help find specific hex codes, such as [htmlcolorcodes.com/](https://htmlcolorcodes.com/).

```{r}
#| label: results-plot4
#| fig.cap: "Bar chart of trust in government by chosen 2020 presidential candidate with colors, labels, error bars, and title"
#| fig.alt: "Bar chart with x-axis of 'VotedPres2020_selection' with labels Biden, Trump and Other. It has y-axis 'pct_trust' with labels 0.00, 0.05, 0.10 and 0.15. The chart is a bar chart with 3 vertical bars. Bar 1 (Biden) has a height of 0.12 and a color of dark blue. Bar 2 (Trump) has a height of 0.17 and a color of very pale blue. Bar 3 (Other) has a height of 0.06 and color of moderate purple. Error bars are added with the Bar 1 (Biden) error ranging from 0.11 to 0.14, Bar 2 (Trump) error ranging from 0.16 to 0.19, and the Bar 3 (Other) error ranging from 0.02 to 0.14."
pfull <-
  anes_des_der %>%
  ggplot(aes(x = VotedPres2020_selection,
             y = pct_trust,
             fill = VotedPres2020_selection)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = pct_trust_low,
                    ymax = pct_trust_upp),
                width = .2) +
  scale_fill_manual(values = c("#0b3954", "#bfd7ea", "#8d6b94")) +
  xlab("Election choice (2020)") +
  ylab("Usually trust the government") +
  scale_y_continuous(labels = scales::percent) +
  guides(fill = "none") +
  labs(title = "Percent of voters who usually trust the government
       by chosen 2020 presidential candidate",
       caption = "Source: American National Election Studies, 2020")

pfull
```