# National Crime Victimization Survey Vignette {#c09-ncvs-vignette}

<!-- author review -->

The United States National Crime Victimization Survey (NCVS) is a household survey sponsored by the Bureau of Justice Statistics (BJS), which collects data on criminal victimization, including characteristics of the crimes, offenders, and victims. Both household and personal crimes include violent and non-violent crimes. The target population of this survey is all people in the United States age 12 and older living in housing units and noninstitutional group quarters.

The NCVS has been ongoing since 1992. An earlier survey, the National Crime Survey, was run from 1972 to 1991 (@ncvs_tech_2016). The survey is administered using a rotating panel. When an address enters the sample, the residents of that address are interviewed every six months for a total of seven interviews. If the initial residents move away from the address during the period, the new residents are included in the survey, and people are not followed when they move. 

NCVS data is publicly available and distributed by Inter-university Consortium for Political and Social Research (ICPSR)^[https://www.icpsr.umich.edu/web/ICPSR/series/95], with data going back to 1992. The vignette in this book will include data from 2021 (@ncvs_data_2021). The NCVS data structure is complicated, and the User's Guide contains examples for analysis in SAS, SUDAAN, SPSS, and Stata, but not R (@ncvs_user_guide). This vignette will adapt those examples for R. 

## Data structure

The data from ICPSR is distributed with five files, each having its unique identifier indicated:

  - Address Record - `YEARQ`, `IDHH`
  - Household Record - `YEARQ`, `IDHH`
  - Person Record - `YEARQ`, `IDHH`, `IDPER`
  - Incident Record - `YEARQ`, `IDHH`, `IDPER`
  - 2021 Collection Year Incident - `YEARQ`, `IDHH`, `IDPER`

We will focus on the household, person, and incident files. From these files, we selected a subset of columns for examples to use in this vignette. Download the complete files at ICPSR^[https://www.icpsr.umich.edu/web/NACJD/studies/38429].

## Survey notation

The NCVS User Guide (@ncvs_user_guide) uses the following notation:

* $i$ represents NCVVS households, identified on the household-level file on the basis of the household identification number `IDHH`.
* $j$ represents NCVS individual respondents within households $i$, identified on the person-level file on the basis of the person identification number `IDPER`.
* $k$ represents reporting periods (i.e., `YEARQ`) for households $i$ and individual respondent $j$.
* $l$ represents victimization records for respondent $j$ in household $i$ and reporting period $k$. Each record on the NCVS incident-level file is associated with a victimization record $l$.
* $D$ represents one or more domain characteristics of interest in the calculation of NCVS estimates. For victimization totals and proportions, domains can be defined on the basis of crime types (e.g., violent crimes, property crimes), characteristics of victims (e.g., age, sex, household income), or characteristics of the victimizations (e.g., victimizations reported to police, victimizations committed with a weapon present). Domains could also be a combination of all of these types of characteristics. For the calculation of victimization rates, domains are defined on the basis of the characteristics of the victims.
* $A_a$ represents the level $a$ of covariate $A$. Covariate $A$ is defined in the calculation of victimization proportions and represents the characteristic for which the analyst wants to obtain the distribution of victimizations in domain $D$.
* $C$ represents the personal or property crime for which we want to obtain a victimization rate.

In this vignette, we will discuss four estimates:

1. *Victimization totals* estimate the number of criminal victimizations with a given characteristic. As demonstrated below, these can be calculated from any of the design objects. The estimated victimization total, $\hat{t}_D$ for domain $D$ is estimated as

$$ \hat{t}_D = \sum_{ijkl \in D} v_{ijkl}$$

where $v_{ijkl}$ is the series-adjusted victimization weight for household $i$, respondent $j$, reporting period $k$, and victimization $l$, that is `WGTVICCY`. 

2. *Victimization proportions* estimate characteristics among victimizations or victims. Victimization proportions are calculated using the incident design object. The estimated victimization proportion for domain $D$ across level $a$ of covariate $A$, $\hat{p}_{A_a,D}$ is 

$$ \hat{p}_{A_a,D} =\frac{\sum_{ijkl \in A_a, D} v_{ijkl}}{\sum_{ijkl \in D} v_{ijkl}}.$$
The numerator is the number of incidents with a particular characteristic in a domain, and the denominator is the number of incidents in a domain.

3. *Victimization rates* are estimates of the number of victimizations per 1,000 persons or households in the population^[BJS publishes victimization rates per 1,000 which are also presented in these examples]. Victimization rates are calculated using the household or person design objects. The estimated victimization rate for crime $C$ in domain $D$ is

$$V\hat{R}_{C,D}= \frac{\sum_{ijkl \in C,D} v_{ijkl}}{\sum_{ijk \in D} w_{ijk}}\times 1000$$
where $w_{ijk}$ is the person weight (`WGTPERCY`) or household weight (`WGTHHCY`) for personal and household crimes, respectively. The numerator is the number of incidents in a domain, and the denominator is the number of persons or households in a domain.

4. Prevalence rates are estimates of the percentage of the population (persons or households) who are victims of a crime. These are estimated using the household or person design objects. The estimated prevalence rate for crime $C$ in domain $D$ is

$$ P\hat{R}_{C, D}= \frac{\sum_{ijk \in {C,D}} I_{ij}w_{ijk}}{\sum_{ijk \in D} w_{ijk}} \times 100$$

where $I_{ij}$ is an indicator that a person or household in domain $D$ was a victim of crime $C$ at any time in the year. The numerator is the number of people who are victims in domain $D$ for crime $C$, and the denominator is the number of people or households in the population.

## Data file preparation

Some work is necessary to prepare the files before analysis. The design variables indicating pseudostratum (`V2117`) and half-sample code (`V2118`) are only included on the household file, so they must be added to the person and incident files for any analysis.

For victimization rates, we need to know the victimization status for both victims and non-victims. The incident file must be summarized and merged onto the household or person files for household-level and person-level crimes, respectively. We begin this vignette by discussing how to create these incident summary files. This is following Section 2.2 of the NCVS User's Guide.

### Preparing files for estimation of victimization rates

Each record on the incident file represents one victimization, which is not the same as one incident. Some victimizations have several instances that make it difficult for the victim to differentiate the details of these incidents, labeled as "series crimes". Appendix A of the User's Guide indicates how to calculate the series weight in other statistical languages.

Here, we adapt that code adapted for R. Essentially, if a victimization is a series crime, its series weight is top-coded at 10 based on the number of actual victimizations. If an incident is a series crime, but the number of occurrences is unknown, the series weight is set to 6. A description of the variables used in this chunk is included in the following table:

|  | Description | Value | Label |
|:---:|:---:|:---:|:---:|
| V4016 | HOW MANY TIMES INCIDENT OCCUR LAST 6 MOS | 1-996 | Number of times |
|  |  | 997 | Don't know |
| V4017 | HOW MANY INCIDENTS | 1 | 1-5 incidents (not a "series") |
|  |  | 2 | 6 or more incidents |
| V4018 | INCIDENTS SIMILAR IN DETAIL | 1 | Similar |
|  |  | 2 | Different (not in a "series") |
| V4019 | ENOUGH DETAIL TO DISTINGUISH INCIDENTS | 1 | Yes (not a "series") |
|  |  | 2 | No (is a "series") |
| WGTVICCY | ADJUSTED VICTIMIZATION WEIGHT |  | Numeric |

```{r}
#| label: ncvs-vign-incfile
#| message: false
#| cache: TRUE
library(tidyverse)
library(srvyr)
library(gt)
library(osfr)
source("helper-fun/helper-functions.R")

inc_in <- read_rds_tsr("ncvs_2021_incident.rds")

inc_series <- inc_in %>%
  mutate(
    series = case_when(
      V4017 %in% c(1, 8) ~ 1,
      V4018 %in% c(2, 8) ~ 1,
      V4019 %in% c(1, 8) ~ 1,
      TRUE ~ 2 # series
    ),
    n10v4016 = case_when(V4016 %in% c(997, 998) ~ NA_real_,
                         V4016 > 10 ~ 10,
                         TRUE ~ V4016),
    serieswgt = case_when(series == 2 & is.na(n10v4016) ~ 6,
                          series == 2 ~ n10v4016,
                          TRUE ~ 1),
    NEWWGT = WGTVICCY * serieswgt
  )
```

The next step in preparing the files for estimation is to create indicators on the victimization file for characteristics of interest. Almost all BJS publications limit the analysis to records where the victimization occurred in the United States, where `V4022` is not equal to 1, and we will do this for all estimates as well.

In the following example, we will create the following indicators:

1. Property crime
    - `V4529` >= 31
    - Variable: `Property`
2. Violent crime
    - `V4529` <= 20
    - Variable: `Violent`
3. Property crime reported to the police
    - `V4529` >= 31 and `V4399`=1
    - Variable: `Property_ReportPolice`
4. Violent crime reported to the police
    - `V4529` < 31 and `V4399`=1
    - Variable: `Violent_ReportPolice`
5. Aggravated assault without a weapon
    - `V4529` in 11:12 and `V4049`=2
    - Variable: `AAST_NoWeap`
6. Aggravated assault with a firearm
    - `V4529` in 11:12 and `V4049`=1 and (`V4051`=1 or `V4052`=1 or `V4050`=7)
    - Variable: `AAST_Firearm`
7. Aggravated assault with a knife or sharp object
    - `V4529` in 11:12 and `V4049`=1 and (`V4053`=1 or `V4054`=1)
    - Variable: `AAST_Knife`
8. Aggravated assault with another type of weapon
    - `V4529` in 11:12 and `V4049`=1 and `V4050`=1 and not firearm or knife
    - Variable: `AAST_Other`

A brief codebook for this section is as follows:

| Variable | Description | Value | Label |
|:---:|:---:|:---:|:---:|
| V4022 | IN WHAT CITY,  TOWN, VILLAGE | 1 | Outside U.S. |
|  |  | 2 | Not inside a city/town/village |
|  |  | 3 | Same city/town/village as present residence |
|  |  | 4 | Different city/town/village as present residence |
|  |  | 5 | Don't know |
|  |  | 6 | Don't know if 2, 4, or 5 |
| V4049 | DID OFFENDER HAVE A WEAPON | 1 | Yes |
|  |  | 2 | No |
|  |  | 3 | Don't know |
| V4050 | WHAT WAS WEAPON | 1 | At least one good entry |
|  |  | 3 | Indicates "Yes-Type Weapon-NA" |
|  |  | 7 | Indicates "Gun Type Unknown" |
|  |  | 8 | No good entry |
| V4051 | HAND GUN | 0 | No |
|  |  | 1 | Yes |
| V4052 | OTHER GUN | 0 | No |
|  |  | 1 | Yes |
| V4053 | KNIFE | 0 | No |
|  |  | 1 | Yes |
| V4399 | REPORTED TO POLICE | 1 | Yes |
|  |  | 2 | No |
|  |  | 3 | Don't know |
| V4529 | TYPE OF CRIME CODE | 01 | Completed rape |
|  |  | 02 | Attempted rape |
|  |  | 03 | Sexual attack with serious assault |
|  |  | 04 | Sexual attack with minor assault |
|  |  | 05 | Completed robbery with injury from serious assault |
|  |  | 06 | Completed robbery with injury from minor assault |
|  |  | 07 | Completed robbery without injury from minor assault |
|  |  | 08 | Attempted robbery with injury from serious assault |
|  |  | 09 | Attempted robbery with injury from minor assault |
|  |  | 10 | Attempted robbery without injury |
|  |  | 11 | Completed aggravated assault with injury |
|  |  | 12 | Attempted aggravated assault with weapon |
|  |  | 13 | Threatened assault with weapon |
|  |  | 14 | Simple assault completed with injury |
|  |  | 15 | Sexual assault without injury |
|  |  | 16 | Unwanted sexual contact without force |
|  |  | 17 | Assault without weapon without injury |
|  |  | 18 | Verbal threat of rape |
|  |  | 19 | Verbal threat of sexual assault |
|  |  | 20 | Verbal threat of assault |
|  |  | 21 | Completed purse snatching |
|  |  | 22 | Attempted purse snatching |
|  |  | 23 | Pocket picking (completed only) |
|  |  | 31 | Completed burglary, forcible entry |
|  |  | 32 | Completed burglary, unlawful entry without force |
|  |  | 33 | Attempted forcible entry |
|  |  | 40 | Completed motor vehicle theft |
|  |  | 41 | Attempted motor vehicle theft |
|  |  | 54 | Completed theft less than $10 |
|  |  | 55 | Completed theft $10 to $49 |
|  |  | 56 | Completed theft $50 to $249 |
|  |  | 57 | Completed theft $250 or greater |
|  |  | 58 | Completed theft value NA |
|  |  | 59 | Attempted theft |

Below, we check the derived variables. This is a good point to pause to look at the output of crosswalks between an original variable and a derived one to check the logic was programmed correctly, and everything ends up in the expected bucket.  

```{r}
#| label: ncvs-vign-inc-inds

inc_ind <- inc_series %>%
  filter(V4022 != 1) %>%
  mutate(
    WeapCat = case_when(
      is.na(V4049) ~ NA_character_,
      V4049 == 2 ~ "NoWeap",
      V4049 == 3 ~ "UnkWeapUse",
      V4050 == 3 ~ "Other",
      V4051 == 1 | V4052 == 1 | V4050 == 7 ~ "Firearm",
      V4053 == 1 | V4054 == 1 ~ "Knife",
      TRUE ~ "Other"
    ),
    V4529_num = parse_number(as.character(V4529)),
    ReportPolice = V4399 == 1,
    Property = V4529_num >= 31,
    Violent = V4529_num <= 20,
    Property_ReportPolice = Property & ReportPolice,
    Violent_ReportPolice = Violent & ReportPolice,
    AAST = V4529_num %in% 11:13,
    AAST_NoWeap = AAST & WeapCat == "NoWeap",
    AAST_Firearm = AAST & WeapCat == "Firearm",
    AAST_Knife = AAST & WeapCat == "Knife",
    AAST_Other = AAST & WeapCat == "Other"
  )

# Checking derived variables
inc_series %>% count(V4022)
inc_ind %>% count(V4022)
inc_ind %>%
  count(WeapCat, V4049, V4050, V4051, V4052, V4052, V4053, V4054)
inc_ind %>% count(V4529, Property, Violent, AAST) %>% print(n = 40)
inc_ind %>% count(ReportPolice, V4399)
inc_ind %>%
  count(AAST,
        WeapCat,
        AAST_NoWeap,
        AAST_Firearm,
        AAST_Knife,
        AAST_Other)
```

After creating indicators of victimization types and characteristics, the file is summarized, and crimes are summed across persons or households by `YEARQ.` Property crimes (i.e., crimes committed against households, such as household burglary or motor vehicle theft) are summed across households, and personal crimes (i.e., crimes committed against an individual, such as assault, robbery, and personal theft) are summed across persons. The indicators are summed using the `serieswgt`, and the variable `WGTVICCY` needs to be retained for later analysis.

```{r}
#| label: ncvs-vign-inc-sum
inc_hh_sums <-
  inc_ind %>%
  filter(V4529_num > 23) %>% # restrict to household crimes
  group_by(YEARQ, IDHH) %>%
  summarize(
    WGTVICCY = WGTVICCY[1],
    across(starts_with("Property"), ~ sum(. * serieswgt),
           .names = "{.col}"),
    .groups = "drop"
  )

inc_pers_sums <-
  inc_ind %>%
  filter(V4529_num <= 23) %>% # restrict to person crimes
  group_by(YEARQ, IDHH, IDPER) %>%
  summarize(WGTVICCY = WGTVICCY[1],
            across(c(
              starts_with("Violent"), starts_with("AAST")
            ),
            ~ sum(. * serieswgt), .names = "{.col}"),
            .groups = "drop")
```

Now, we merge the victimization summary files into the appropriate files. For any record on the household or person file that is not on the victimization file, the victimization counts are set to 0 after merging. In this step, we will also create the victimization adjustment factor. See 2.2.4 in the User's Guide for details of why this adjustment is created. It is calculated as follows:

$$ A_{ijk}=\frac{v_{ijk}}{w_{ijk}}$$

where $w_{ijk}$ is the person weight (`WGTPERCY`) for personal crimes or the household weight (`WGTHHCY`) for household crimes, and $v_{ijk}$ is the victimization weight (`WGTVICCY`) for household $i$, respondent $j$, in reporting period $k$. The adjustment factor is set to 0 if no incidents are reported.

```{r}
#| label: ncvs-vign-merge-inc-sum
#| cache: TRUE
hh_in <- read_rds_tsr("ncvs_2021_household.rds")

pers_in <- read_rds_tsr("ncvs_2021_person.rds")

# Set up a list of 0s for each crime type/characteristic to replace NA's
hh_z_list <- rep(0, ncol(inc_hh_sums) - 3) %>% as.list() %>%
  setNames(names(inc_hh_sums)[-(1:3)])
pers_z_list <- rep(0, ncol(inc_pers_sums) - 4) %>% as.list() %>%
  setNames(names(inc_pers_sums)[-(1:4)])

hh_vsum <- hh_in %>%
  full_join(inc_hh_sums, by = c("YEARQ", "IDHH")) %>%
  replace_na(hh_z_list) %>%
  mutate(ADJINC_WT = if_else(is.na(WGTVICCY), 0, WGTVICCY / WGTHHCY))

pers_vsum <- pers_in %>%
  full_join(inc_pers_sums, by = c("YEARQ", "IDHH", "IDPER")) %>%
  replace_na(pers_z_list) %>%
  mutate(ADJINC_WT = if_else(is.na(WGTVICCY), 0, WGTVICCY / WGTPERCY))
```

### Derived demographic variables

A final step in file preparation for the household and person files is creating any derived variables on the household and person files, such as income categories or age categories, for subgroup analysis. We can do this step before or after merging the victimization counts.

For the household file, we create categories for tenure (rental status), urbanicity, income, place size, and region. For the person file, we create categories for sex, race/Hispanic origin, age categories, and marital status. We also merge the household demographics to the person file as well as the design variables (`V2117` and `V2118`).

Brief codebook for household variables:

|Variable|Description|Value|Label|
|---|---|---|---|
|V2015|TENURE|1|Owned or being bought|
|||2|Rented for cash|
|||3|No cash rent|
|SC214A|HOUSEHOLD INCOME|01|Less than $5,000|
|||02|$5,000 to $7,499|
|||03|$7,500 to $9,999|
|||04|$10,000 to $12,499|
|||05|$12,500 to $14,999|
|||06|$15,000 to $17,499|
|||07|$17,500 to $19,999|
|||08|$20,000 to $24,999|
|||09|$25,000 to $29,999|
|||10|$30,000 to $34,999|
|||11|$35,000 to $39,999|
|||12|$40,000 to $49,999|
|||13|$50,000 to $74,999|
|||14|$75,000 and over|
|||15|$75,000 to $99,999|
|||16|$100,000-$149,999|
|||17|$150,000-$199,999|
|||18|$200,000 or more|
|V2126B|PLACE SIZE CODE|00|Not in a place|
|||13|Under 10,000|
|||16|10,000-49,999|
|||17|50,000-99,999|
|||18|100,000-249,999|
|||19|250,000-499,999|
|||20|500,000-999,999|
|||21|1,000,000-2,499,999|
|||22|2,500,000-4,999,999|
|||23|5,000,000 or more|
|V2127B|REGION|1|Northeast|
|||2|Midwest|
|||3|South|
|||4|West|
|V2143|URBANICITY|1|Urban|
|||2|Suburban|
|||3|Rural|

Brief codebook for person variables:

|Variable|Description|Value|Label| 
|---|---|---|---|
|V3014|AGE||12 through 90
|V3015|MARITAL STATUS (Current)|1|Married|
|||2|Widowed|
|||3|Divorced|
|||4|Separated|
|||5|Never married|
|V3018|sex|1|Male|
|||2|Female|
|V3023A|RACE|01|White only|
|||02|Black only|
|||03|American Indian, Alaska native only|
|||04|Asian only|
|||05|Hawaiian/Pacific Islander only|
|||06|White-Black|
|||07|White-American Indian|
|||08|White-Asian|
|||09|White-Hawaiian|
|||10|Black-American Indian|
|||11|Black-Asian|
|||12|Black-Hawaiian/Pacific Islander|
|||13|American Indian-Asian|
|||14|Asian-Hawaiian/Pacific Islander|
|||15|White-Black-American Indian|
|||16|White-Black-Asian|
|||17|White-American Indian-Asian|
|||18|White-Asian-Hawaiian|
|||19|2 or 3 races|
|||20|4 or 5 races|
|V3024|HISPANIC ORIGIN|1|Yes|
|||2|No|

```{r}
#| label: ncvs-vign-hh-pers-der
hh_vsum_der <- hh_vsum %>%
  mutate(
    Tenure = factor(
      case_when(V2015 == 1 ~ "Owned", !is.na(V2015) ~ "Rented"),
      levels = c("Owned", "Rented")
    ),
    Urbanicity = factor(
      case_when(V2143 == 1 ~ "Urban",
                V2143 == 2 ~ "Suburban",
                V2143 == 3 ~ "Rural"),
      levels = c("Urban", "Suburban", "Rural")
    ),
    SC214A_num = as.numeric(as.character(SC214A)),
    Income = case_when(
      SC214A_num <= 8 ~ "Less than $25,000",
      SC214A_num <= 12 ~ "$25,000-49,999",
      SC214A_num <= 15 ~ "$50,000-99,999",
      SC214A_num <= 17 ~ "$100,000-199,999",
      SC214A_num <= 18 ~ "$200,000 or more"
    ),
    Income = fct_reorder(Income, SC214A_num, .na_rm = FALSE),
    PlaceSize = case_match(
      as.numeric(as.character(V2126B)),
      0 ~ "Not in a place",
      13 ~ "Under 10,000",
      16 ~ "10,000-49,999",
      17 ~ "50,000-99,999",
      18 ~ "100,000-249,999",
      19 ~ "250,000-499,999",
      20 ~ "500,000-999,999",
      c(21, 22, 23) ~ "1,000,000 or more"
    ),
    PlaceSize = fct_reorder(PlaceSize, as.numeric(V2126B)),
    Region = case_match(
      as.numeric(V2127B),
      1 ~ "Northeast",
      2 ~ "Midwest",
      3 ~ "South",
      4 ~ "West"
    ),
    Region = fct_reorder(Region, as.numeric(V2127B))
  )

hh_vsum_der %>% count(Tenure, V2015)
hh_vsum_der %>% count(Urbanicity, V2143)
hh_vsum_der %>% count(Income, SC214A)
hh_vsum_der %>% count(PlaceSize, V2126B)
hh_vsum_der %>% count(Region, V2127B)

hh_vsum_slim <- hh_vsum_der %>%
  select(YEARQ:V2118,
         WGTVICCY:ADJINC_WT,
         Tenure,
         Urbanicity,
         Income,
         PlaceSize,
         Region)


NHOPI <-
  "Native Hawaiian or Other Pacific Islander" # made for usage later
pers_vsum_der <- pers_vsum %>%
  mutate(
    Sex = factor(case_when(V3018 == 1 ~ "Male",
                           V3018 == 2 ~ "Female")),
    RaceHispOrigin = factor(
      case_when(
        V3024 == 1 ~ "Hispanic",
        V3023A == 1 ~ "White",
        V3023A == 2 ~ "Black",
        V3023A == 4 ~ "Asian",
        V3023A == 5 ~ NHOPI,
        TRUE ~ "Other"
      ),
      levels = c("White", "Black", "Hispanic", "Asian", NHOPI, "Other")
    ),
    V3014_num = as.numeric(as.character(V3014)),
    AgeGroup = case_when(
      V3014_num <= 17 ~ "12-17",
      V3014_num <= 24 ~ "18-24",
      V3014_num <= 34 ~ "25-34",
      V3014_num <= 49 ~ "35-49",
      V3014_num <= 64 ~ "50-64",
      V3014_num <= 90 ~ "65 or older",
    ),
    AgeGroup = fct_reorder(AgeGroup, V3014_num),
    MaritalStatus = factor(
      case_when(
        V3015 == 1 ~ "Married",
        V3015 == 2 ~ "Widowed",
        V3015 == 3 ~ "Divorced",
        V3015 == 4 ~ "Separated",
        V3015 == 5 ~ "Never married"
      ),
      levels = c("Never married", "Married", "Widowed",
                 "Divorced", "Separated")
    )
  ) %>%
  left_join(select(hh_vsum_slim, YEARQ, IDHH, V2117, V2118, Tenure:Region),
            by = c("YEARQ", "IDHH"))

pers_vsum_der %>% count(Sex, V3018)
pers_vsum_der %>% count(RaceHispOrigin, V3024)
pers_vsum_der %>%
  filter(RaceHispOrigin != "Hispanic" | is.na(RaceHispOrigin)) %>%
  count(RaceHispOrigin, V3023A)
pers_vsum_der %>% group_by(AgeGroup) %>%
  summarize(minAge = min(V3014),
            maxAge = max(V3014),
            .groups = "drop")
pers_vsum_der %>% count(MaritalStatus, V3015)

pers_vsum_slim <- pers_vsum_der %>%
  select(YEARQ:WGTPERCY, WGTVICCY:ADJINC_WT, Sex:Region)
```

The tibbles `hh_vsum_slim` and `pers_vsum_slim` can now be used to create design objects and calculate crime rate estimates. To calculate estimates about types of crime, such as what percentage of violent crimes are reported to the police, we must use the incident file. The incident file is not guaranteed to have every pseudostratum and half-sample code, so dummy records are created to append before estimation. Finally, demographic variables are merged onto the incident tibble.

```{r}
#| label: ncvs-vign-inc-analysis
dummy_records <- hh_vsum_slim %>%
  distinct(V2117, V2118) %>%
  mutate(Dummy = 1,
         WGTVICCY = 1,
         NEWWGT = 1)

inc_analysis <- inc_ind %>%
  mutate(Dummy = 0) %>%
  left_join(select(pers_vsum_slim, YEARQ, IDHH, IDPER, Sex:Region),
            by = c("YEARQ", "IDHH", "IDPER")) %>%
  bind_rows(dummy_records) %>%
  select(YEARQ:IDPER,
         WGTVICCY,
         NEWWGT,
         V4529,
         WeapCat,
         ReportPolice,
         Property:Region)
```

## Survey design objects

All the data prep above is necessary to prepare the data for survey analysis. At this point, we can create the design objects and finally begin analysis. We will create three design objects for different types of analysis as they depend on which type of estimate we are creating. For the incident data, the weight of analysis is `NEWWGT`, which we constructed previously. The household and person data use `WGTHHCY` and `WGTPERCY`, respectively. For all analyses, `V2117` is the strata variable, and `V2118` is the cluster variable for analysis.

```{r}
#| label: ncvs-vign-desobj

inc_des <- inc_analysis %>%
  as_survey(
    weight = NEWWGT,
    strata = V2117,
    ids = V2118,
    nest = TRUE
  )

hh_des <- hh_vsum_slim %>%
  as_survey(
    weight = WGTHHCY,
    strata = V2117,
    ids = V2118,
    nest = TRUE
  )

pers_des <- pers_vsum_slim %>%
  as_survey(
    weight = WGTPERCY,
    strata = V2117,
    ids = V2118,
    nest = TRUE
  )
```

## Calculating estimates

Now that we have prepared our data, we can calculate our estimates. As a reminder, those are:

1. *Victimization totals* estimate the number of criminal victimizations with a given characteristic.

2. *Victimization proportions* estimate characteristics among victimizations or victims.

3. *Victimization rates* are estimates of the number of victimizations per 1,000 persons or households in the population.

4. Prevalence rates are estimates of the percentage of the population (persons or households) who are victims of a crime.

### Victimization totals

There are two ways to calculate victimization totals. Using the incident design object is the most straightforward method, but the person and household design objects can be used as well if the adjustment factor is incorporated. In the example below, the total number of property and violent victimizations are first calculated using the incident file and then using the household and person design objects. The incident file is smaller, and thus, estimation is faster using that file, but the estimates will be the same as illustrated below:

```{r}
#| label: ncvs-vign-victot-examp
inc_des %>%
  summarize(
    Property_Vzn = survey_total(Property, na.rm = TRUE),
    Violent_Vzn = survey_total(Violent, na.rm = TRUE)
  )

hh_des %>%
  summarize(Property_Vzn = survey_total(Property * ADJINC_WT, na.rm = TRUE))

pers_des %>%
  summarize(Violent_Vzn = survey_total(Violent * ADJINC_WT, na.rm = TRUE))
```

### Victimization proportions

Victimization proportions are proportions describing features of a victimization. The key here is that these are questions among victimizations, not among the population. These types of estimates can only be calculated using the incident design object. These include questions such as:

- What proportion of property victimizations are reported to the police?
- What proportion of violent victimizations are in urban areas?

```{r}
#| label: ncvs-vign-vic-prop
inc_des %>%
  filter(Property) %>%
  group_by(ReportPolice) %>%
  summarize(Pct = survey_prop(na.rm = TRUE) * 100)

inc_des %>%
  filter(Violent) %>%
  group_by(Urbanicity) %>%
  summarize(Pct = survey_prop(na.rm = TRUE) * 100)
```

### Victimization rates

Victimization rates measure the number of victimizations per population and are not an estimate of the proportion of households or persons who are victimized, which is a prevalence rate described in the next section. Victimization rates are estimated using the household or person design objects depending on the type of crime, and the adjustment factor must be incorporated. We return to the example of property and violent victimizations used in the example for victimization totals. In the following example, the property victimization totals are calculated as above, as well as the property victimization rate, using `survey_mean()`, and the population size using `survey_total()`.

```{r}
#| label: ncvs-vign-vic-rate
hh_des %>%
  summarize(
    Property_Vzn = survey_total(Property * ADJINC_WT, na.rm = TRUE),
    Property_Rate = survey_mean(Property * ADJINC_WT * 1000,
                                na.rm = TRUE),
    PopSize = survey_total(1, vartype = NULL)
  )
```

Victimization rates can also be calculated for particular characteristics of the victimization. In the following example, the rate of aggravated assault with no weapon, with a firearm, with a knife, and with another weapon.

```{r}
#| label: ncvs-vign-pers-rates-char
pers_des %>%
  summarize(across(
    starts_with("AAST_"),
    ~ survey_mean(. * ADJINC_WT * 1000, na.rm = TRUE)
  ))
```

A common desire is to calculate victimization rates by several characteristics. For example, we may want to calculate the violent victimization rate and aggravated assault rate by sex, race/Hispanic origin, age group, marital status, and household income. This requires a `group_by()` statement for each categorization separately. Thus, we make a function to do this and then use `map_df()` from the {purrr} package to loop through the variables. Finally, the {gt} package is used to make a publishable table.

```{r}
#| ncvs-vign-rates-demo
pers_est_by <- function(byvar) {
  pers_des %>%
    rename(Level := {
      {
        byvar
      }
    }) %>%
    filter(!is.na(Level)) %>%
    group_by(Level) %>%
    summarize(
      Violent = survey_mean(Violent * ADJINC_WT * 1000, na.rm = TRUE),
      AAST = survey_mean(AAST * ADJINC_WT * 1000, na.rm = TRUE)
    ) %>%
    mutate(
      Variable = byvar,
      LevelNum = as.numeric(Level),
      Level = as.character(Level)
    ) %>%
    select(Variable, Level, LevelNum, everything())
}

pers_est_df <-
  c("Sex", "RaceHispOrigin", "AgeGroup", "MaritalStatus", "Income") %>%
  map_df(pers_est_by)

pers_est_df %>%
  mutate(
    Variable = case_when(
      Variable == "RaceHispOrigin" ~ "Race/Hispanic origin",
      Variable == "MaritalStatus" ~ "Marital status",
      Variable == "AgeGroup" ~ "Age",
      TRUE ~ Variable
    )
  ) %>%
  select(-LevelNum) %>%
  group_by(Variable) %>%
  gt(rowname_col = "Level") %>%
  tab_spanner(
    label = "Violent crime",
    id = "viol_span",
    columns = c("Violent", "Violent_se")
  ) %>%
  tab_spanner(label = "Aggravated assault",
              columns = c("AAST", "AAST_se")) %>%
  cols_label(
    Violent = "Rate",
    Violent_se = "SE",
    AAST = "Rate",
    AAST_se = "SE",
  ) %>%
  fmt_number(
    columns = c("Violent", "Violent_se", "AAST", "AAST_se"),
    decimals = 1
  ) %>%
  tab_footnote(
    footnote = "Includes rape or sexual assault, robbery,
    aggravated assault, and simple assault.",
    locations = cells_column_spanners(spanners = "viol_span")
  ) %>%
  tab_footnote(footnote = "Excludes persons of Hispanic origin",
               locations = cells_stub(rows = Level != "Hispanic")) %>% 
  tab_footnote(
    footnote = "Inlcudes persons who identified as
    Native Hawaiian or Other Pacific Islander only.",
    locations = cells_stub(rows = Level == NHOPI)
  ) %>%
  tab_footnote(
    footnote = "Inlcudes persons who identified as American Indian or
    Alaska Native only or as two or more races.",
    locations = cells_stub(rows = Level == "Other")
  ) %>%
  tab_source_note(source_note = "Note: Rates per 1,000 persons age 12 or older.") %>%
  tab_source_note(source_note = "Source: Bureau of Justice Statistics,
                  National Crime Victimization Survey, 2021.") %>%
  tab_stubhead(label = "Victim demographic") %>%
  tab_header(title = "Rate and standard error of violent victimization,
             by type of crime and demographic characteristics, 2021")
```

### Prevalence rates

Prevalence rates are different from victimization rates as the numerator is the number of people or households who are victimized rather than the number of victimizations. To calculate the prevalence rates, another summary of the data must be done, which is calculating an indicator for whether a person or household is a victim of a particular crime at any point in the year. Below is an example of calculating first the indicator and then the prevalence rate of violent crime and aggravated assault.

```{r}
#| label: ncvs-vign-prevexamp
pers_prev_des <-
  pers_des %>%
  mutate(Year = floor(YEARQ)) %>%
  group_by(Year, IDHH, IDPER) %>%
  mutate(Violent_Ind = sum(Violent) > 0,
         AAST_Ind = sum(AAST) > 0) %>%
  ungroup()

pers_prev_ests <- pers_prev_des %>%
  summarize(Violent_Prev = survey_mean(Violent_Ind * 100),
            AAST_Prev = survey_mean(AAST_Ind * 100))

pers_prev_ests
```

In the example above, the indicator is multiplied by 100 to return a percentage rather than a proportion. In 2021, we estimate that `r round(pers_prev_ests$Violent_Prev, 2)`% of people aged 12 and older were a victim of violent crime in the United States, and `r round(pers_prev_ests$AAST_Prev, 2)`% were victims of aggravated assault.

## Exercises

1. What proportion of completed motor vehicle thefts are not reported to the police? Hint: Use the codebook to look at the definition of Type of Crime (V4529).

```{r}
#| label: ncvs-vign-des-sol1
ans1 <- inc_des %>%
  filter(str_detect(V4529, "40|41")) %>%
  group_by(ReportPolice) %>%
  summarize(Pct = survey_prop(na.rm = TRUE) * 100)
ans1
ans1 %>% filter(!ReportPolice) %>% pull(Pct)
```

2. How many violent crimes occur in each region?

```{r}
#| label: ncvs-vign-des-sol2
inc_des %>%
  filter(Violent) %>%
  survey_count(Region)
```

3. What is the property victimization rate among each income level?

```{r}
#| label: ncvs-vign-des-sol3
hh_des %>%
  group_by(Income) %>%
  summarize(Property_Rate = survey_mean(Property * ADJINC_WT * 1000, na.rm =
                                          TRUE))
```
