\cleardoublepage 

# (APPENDIX) Appendices {-}

# ANES Derived Variable Codebook {#anes-cb}

```{r}
#| label: anes-cb-setup
#| echo: FALSE
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(janitor)
library(kableExtra)
library(knitr)

anes_2020 <- anes_2020 %>% haven::zap_labels()

cb_count <- function(dat, ...){
  t <- dat %>%
    count(...) %>%
    mutate(`Unweighted Freq` = n / sum(n)) %>%
    janitor::adorn_totals(where="row", fill="-", na.rm=TRUE, name="Total", n, `Unweighted Freq`) %>%
    mutate(`Unweighted Freq`= round(`Unweighted Freq`, 3)) %>%
    kbl(position="H")
  
   if (knitr:::is_html_output()){
    t %>% kable_minimal()
  } else{
    t 
  }
}

cb_continuous <- function(dat, var){
  t <- dat %>%
    summarize(
      `N Missing`=sum(is.na({{var}})),
      Minimum = min({{var}}, na.rm = TRUE),
      Median = median({{var}}, na.rm = TRUE),
      Maximum = max({{var}}, na.rm = TRUE)) %>%
    kbl(position="H") 
  
  if (knitr:::is_html_output()){
    t %>% kable_minimal()
  } else{
    t 
  }
  
}


```

The full codebook with the original variables is available at [https://electionstudies.org/wp-content/uploads/2022/02/anes_timeseries_2020_userguidecodebook_20220210.pdf](https://electionstudies.org/wp-content/uploads/2022/02/anes_timeseries_2020_userguidecodebook_20220210.pdf)

## Weighting variables

#### V200010b {-}

Full sample post-election weight.

#### V200010d {-}

Full sample variance stratum.

#### V200010c {-}

Full sample variance unit.

## Additional variables

#### InterviewMode {-}

MODE OF INTERVIEW: PRE-ELECTION INTERVIEW

```{r}
#| label: anes-cb-interviewmode
#| echo: FALSE
anes_2020 %>%
   cb_count(V200002, InterviewMode)
```
Variables used: V200002

#### Age {-}

RESPONDENT AGE

```{r}
#| label: anes-cb-age
#| echo: FALSE
anes_2020 %>%
   cb_count(AgeGroup)
```

Variables used: V201507x

#### Gender {-}

PRE: WHAT IS YOUR (R) SEX? [REVISED]

What is your sex?

```{r}
#| label: anes-cb-gender
#| echo: FALSE
anes_2020 %>%
   cb_count(V201600, Gender)
```

Variables used: V201600

#### RaceEth {-}

PRE: SUMMARY: R SELF-IDENTIFIED RACE/ETHNICITY

```{r}
#| label: anes-cb-raceeth
#| echo: FALSE
anes_2020 %>%
   cb_count(V201549x, RaceEth) 
```

Variables used: V201549x

#### PartyID {-}

PRE: SUMMARY: PARTY ID

```{r}
#| label: anes-cb-partyid
#| echo: FALSE
anes_2020 %>%
   cb_count(V201231x, PartyID) 
```

Variables used: V201231x

#### Education {-}

What is the highest level of school you have completed or the
highest degree you have received?

```{r}
#| label: anes-cb-education
#| echo: FALSE
anes_2020 %>%
   cb_count(V201510, Education) 
```

Variables used: V201510

#### Income {-}

PRE: SUMMARY: TOTAL (FAMILY) INCOME

```{r}
#| label: anes-cb-income
#| echo: FALSE
anes_2020 %>%
   cb_count(V201617x, Income)
```

```{r}
#| label: anes-cb-income7
#| echo: FALSE
anes_2020 %>%
   cb_count(V201617x, Income7)
```

Variables used: V201617x

#### CampaignInterest {-}

PRE: HOW INTERESTED IN FOLLOWING CAMPAIGNS

Some people don’t pay much attention to political campaigns. How about you? Would you say that you have been very much interested, somewhat interested or not much interested in the political campaigns so far this year?

```{r}
#| label: anes-cb-campaigninterest
#| echo: FALSE
anes_2020 %>% 
   cb_count(V201006, CampaignInterest) 
```

Variables used: V201006

#### TrustGovernment {-}

PRE: HOW OFTEN TRUST GOVERNMENT IN WASHINGTON TO DO WHAT IS RIGHT [REVISED]

How often can you trust the federal government in Washington to do what is right?

```{r}
#| label: anes-cb-trustgovernment
#| echo: FALSE
anes_2020 %>%
   cb_count(V201233, TrustGovernment) 
```

Variables used: V201233

#### TrustPeople {-}

PRE: HOW OFTEN CAN PEOPLE BE TRUSTED

Generally speaking, how often can you trust other people?

```{r}
#| label: anes-cb-trustpeople
#| echo: FALSE
anes_2020 %>%
   cb_count(V201237, TrustPeople) 
```

Variables used: V201237

#### VotedPres2016 {-}

PRE: DID R VOTE FOR PRESIDENT IN 2016

Four years ago, in 2016, Hillary Clinton ran on the Democratic ticket against Donald Trump for the Republicans. Do you remember for sure whether or not you voted in that election?

*Revised:* Four years ago, in 2016, Hillary Clinton ran on the Democratic ticket against Donald Trump for the Republicans. We talk to many people who tell us they did not vote. And we talk to a few people who tell us they did vote, who really did not. We can tell they did not vote by checking with official government records. What about you? If we check the official government voter records, will they show that you voted in the 2016 presidential election, or that you did not vote in that election?

```{r}
#| label: anes-cb-votedpres2016
#| echo: FALSE
anes_2020 %>%
   cb_count(V201101, V201102, VotedPres2016) 
```

Variables used: V201101, V201102

#### VotedPres2016_selection {-}

PRE: RECALL OF LAST (2016) PRESIDENTIAL VOTE CHOICE

Which one did you vote for?

```{r}
#| label: anes-cb-votedpres2016-selection
#| echo: FALSE
anes_2020 %>%
   cb_count(V201103, VotedPres2016_selection) 
```

Variables used: V201103

#### VotedPres2020 {-}

PRE-POST: SUMMARY: VOTER TURNOUT IN 2020

```{r}
#| label: anes-cb-votedpres2020
#| echo: FALSE
anes_2020 %>%
   cb_count(V202109x, VotedPres2020)
```

Variables used: V202109x

#### VotedPres2020_selection {-}

POST: FOR WHOM DID R VOTE FOR PRESIDENT

```{r}
#| label: anes-cb-votedpres2020-selection
#| echo: FALSE
anes_2020 %>%
   cb_count(V202073, VotedPres2020_selection)
```

Variables used: V202073

#### EarlyVote2020 / VotedPres2020 {-}

PRE: SUMMARY: REGISTRATION AND EARLY VOTE STATUS

```{r}
#| label: anes-cb-earlyvote2020
#| echo: FALSE
anes_2020 %>%
   cb_count(V201025x, V202109x, VotedPres2020, EarlyVote2020)
```

Variables used: V201025x, V202109x
