# RECS Derived Variable Codebook

```{r}
#| label: recs-cb-cb-setup
#| echo: FALSE
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(janitor)
library(kableExtra)
library(knitr)

recs <- recs_in %>% haven::zap_labels()
```

The full codebook with the original variables is available at [https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata](https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata).

## Weighting variables

#### DOEID {-}

Unique identifier for each respondent

#### NWEIGHT {-}

Final sample weight

#### BRRWT {-}

Replicate weights

## Categorical 

#### Region {-}

```{r}
#| label: recs-cb-region
#| echo: FALSE
recs %>%
  count(Region) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```
Variables used: REGIONC

#### Division {-}

```{r}
#| label: recs-cb-division
#| echo: FALSE
recs %>%
  count(Division) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```
Variables used: DIVISION

#### MSAStatus {-}

```{r}
#| label: recs-cb-msastatus
#| echo: FALSE
recs %>%
  count(MSAStatus) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```
Variables used: METROMICRO

#### Urbanicity {-}

```{r}
#| label: recs-cb-urbanicity
#| echo: FALSE
recs %>%
  count(Urbanicity) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: UATYP10

#### HousingUnitType {-}

```{r}
#| label: recs-cb-housingunittype
#| echo: FALSE
recs %>%
  count(HousingUnitType) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TYPEHUQ

#### YearMade {-}

```{r}
#| label: recs-cb-yearmade
#| echo: FALSE
recs %>%
  count(YearMade) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: YEARMADERANGE

#### SpaceHeatingUsed {-}

```{r}
#| label: recs-cb-spaceheadingused
#| echo: FALSE
recs %>%
  count(SpaceHeatingUsed) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: HEATHOME

#### HeatingBehavior {-}

```{r}
#| label: recs-cb-heatingbehavior
#| echo: FALSE
recs %>%
  count(HeatingBehavior) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: EQUIPMUSE

#### ACUsed {-}

```{r}
#| label: recs-cb-acused
#| echo: FALSE
recs %>%
  count(ACUsed) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: AIRCOND

#### ACBehavior {-}

```{r}
#| label: recs-cb-acbehavior
#| echo: FALSE
recs %>%
  count(ACBehavior) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: USECENAC

#### ClimateRegion_BA {-}

```{r}
#| label: recs-cb-climateregion-ba
#| echo: FALSE
recs %>%
  count(ClimateRegion_BA) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: CLIMATE_REGION_PUB

#### ClimateRegion_IECC {-}

```{r}
#| label: recs-cb-climateregion-iecc
#| echo: FALSE
recs %>%
  count(ClimateRegion_IECC) %>%
  mutate(`Unweighted Freq` = round(n / sum(n), 3)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: IECC_CLIMATE_PUB

## Continuous

#### WinterTempDay {-}

```{r}
#| label: recs-cb-wintertempday
#| echo: FALSE
recs %>%
  summarize(Minimum = min(WinterTempDay, na.rm = TRUE),
            Median = median(WinterTempDay, na.rm = TRUE),
            Maximum = max(WinterTempDay, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPHOME

#### WinterTempAway {-}

```{r}
#| label: recs-cb-wintertempaway
#| echo: FALSE
recs %>%
  summarize(Minimum = min(WinterTempAway, na.rm = TRUE),
            Median = median(WinterTempAway, na.rm = TRUE),
            Maximum = max(WinterTempAway, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPGONE

#### WinterTempNight {-}

```{r}
#| label: recs-cb-wintertempnight
#| echo: FALSE
recs %>%
  summarize(Minimum = min(WinterTempNight, na.rm = TRUE),
            Median = median(WinterTempNight, na.rm = TRUE),
            Maximum = max(WinterTempNight, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPNITE

#### SummerTempDay {-}

```{r}
#| label: recs-cb-summertempday
#| echo: FALSE
recs %>%
  summarize(Minimum = min(SummerTempDay, na.rm = TRUE),
            Median = median(SummerTempDay, na.rm = TRUE),
            Maximum = max(SummerTempDay, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPHOMEAC

#### SummerTempAway {-}

```{r}
#| label: recs-cb-summertempaway
#| echo: FALSE
recs %>%
  summarize(Minimum = min(SummerTempAway, na.rm = TRUE),
            Median = median(SummerTempAway, na.rm = TRUE),
            Maximum = max(SummerTempAway, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPGONEAC

#### SummerTempNight {-}

```{r}
#| label: recs-cb-summertempnight
#| echo: FALSE
recs %>%
  summarize(Minimum = min(SummerTempNight, na.rm = TRUE),
            Median = median(SummerTempNight, na.rm = TRUE),
            Maximum = max(SummerTempNight, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

Variables used: TEMPNITEAC

#### TOTCSQFT {-}

Total cooled square footage

```{r}
#| label: recs-cb-totcsqft
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTCSQFT, na.rm = TRUE),
            Median = median(TOTCSQFT, na.rm = TRUE),
            Maximum = max(TOTCSQFT, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTHSQFT {-}

Total heated square footage

```{r}
#| label: recs-cb-tothsqft
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTHSQFT, na.rm = TRUE),
            Median = median(TOTHSQFT, na.rm = TRUE),
            Maximum = max(TOTHSQFT, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTSQFT_EN {-}

Total square footage (used for publication)

```{r}
#| label: recs-cb-totsqft
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTSQFT_EN, na.rm = TRUE),
            Median = median(TOTSQFT_EN, na.rm = TRUE),
            Maximum = max(TOTSQFT_EN, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTUCSQFT {-}

Total uncooled square footage

```{r}
#| label: recs-cb-totucsqft
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTUCSQFT, na.rm = TRUE),
            Median = median(TOTUCSQFT, na.rm = TRUE),
            Maximum = max(TOTUCSQFT, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTUSQFT {-}

Total unheated square footage

```{r}
#| label: recs-cb-totusqft
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTUSQFT, na.rm = TRUE),
            Median = median(TOTUSQFT, na.rm = TRUE),
            Maximum = max(TOTUSQFT, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### CDD30YR {-}

Cooling degree days, 30-year average 1981-2010, base temperature 65F

```{r}
#| label: recs-cb-cdd30yr
#| echo: FALSE
recs %>%
  summarize(Minimum = min(CDD30YR, na.rm = TRUE),
            Median = median(CDD30YR, na.rm = TRUE),
            Maximum = max(CDD30YR, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### CDD65 {-}

Cooling degree days in 2015, base temperature 65F

```{r}
#| label: recs-cb-cdd65
#| echo: FALSE
recs %>%
  summarize(Minimum = min(CDD65, na.rm = TRUE),
            Median = median(CDD65, na.rm = TRUE),
            Maximum = max(CDD65, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### CDD80 {-}

Cooling degree days in 2015, base temperature 80F (used for garage cooling load estimation only)

```{r}
#| label: recs-cb-cdd80
#| echo: FALSE
recs %>%
  summarize(Minimum = min(CDD80, na.rm = TRUE),
            Median = median(CDD80, na.rm = TRUE),
            Maximum = max(CDD80, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### HDD30YR {-}

Heating degree days, 30-year average 1981-2010, base temperature 65F

```{r}
#| label: recs-cb-hdd30yr
#| echo: FALSE
recs %>%
  summarize(Minimum = min(HDD30YR, na.rm = TRUE),
            Median = median(HDD30YR, na.rm = TRUE),
            Maximum = max(HDD30YR, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### HDD65 {-}

Heating degree days in 2015, base temperature 65F

```{r}
#| label: recs-cb-hdd65
#| echo: FALSE
recs %>%
  summarize(Minimum = min(HDD65, na.rm = TRUE),
            Median = median(HDD65, na.rm = TRUE),
            Maximum = max(HDD65, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### HDD50 {-}

Heating degree days in 2015, base temperature 50F (used for garage heating load estimation only)

```{r}
#| label: recs-cb-hdd50
#| echo: FALSE
recs %>%
  summarize(Minimum = min(HDD50, na.rm = TRUE),
            Median = median(HDD50, na.rm = TRUE),
            Maximum = max(HDD50, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### GNDHDD65 {-}

Heating degree days of ground temperature in 2015, base temperature 65F

```{r}
#| label: recs-cb-gndhdd65
#| echo: FALSE
recs %>%
  summarize(Minimum = min(GNDHDD65, na.rm = TRUE),
            Median = median(GNDHDD65, na.rm = TRUE),
            Maximum = max(GNDHDD65, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTUEL {-}

Total site electricity usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btuel
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTUEL, na.rm = TRUE),
            Median = median(BTUEL, na.rm = TRUE),
            Maximum = max(BTUEL, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### DOLLAREL {-}

Total electricity cost, in dollars, 2015

```{r}
#| label: recs-cb-dollarel
#| echo: FALSE
recs %>%
  summarize(Minimum = min(DOLLAREL, na.rm = TRUE),
            Median = median(DOLLAREL, na.rm = TRUE),
            Maximum = max(DOLLAREL, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTUNG {-}

Total natural gas usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btung
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTUNG, na.rm = TRUE),
            Median = median(BTUNG, na.rm = TRUE),
            Maximum = max(BTUNG, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### DOLLARNG {-}

Total natural gas cost, in dollars, 2015

```{r}
#| label: recs-cb-dollarng
#| echo: FALSE
recs %>%
  summarize(Minimum = min(DOLLARNG, na.rm = TRUE),
            Median = median(DOLLARNG, na.rmx = TRUE),
            Maximum = max(DOLLARNG, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTULP {-}

Total propane usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btulp
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTULP, na.rm = TRUE),
            Median = median(BTULP, na.rm = TRUE),
            Maximum = max(BTULP, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### DOLLARLP {-}

Total cost of propane, in dollars, 2015

```{r}
#| label: recs-cb-dollarlp
#| echo: FALSE
recs %>%
  summarize(Minimum = min(DOLLARLP, na.rm = TRUE),
            Median = median(DOLLARLP, na.rm = TRUE),
            Maximum = max(DOLLARLP, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTUFO {-}

Total fuel oil/kerosene usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btufo
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTUFO, na.rm = TRUE),
            Median = median(BTUFO, na.rm = TRUE),
            Maximum = max(BTUFO, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### DOLLARFO {-}

Total cost of fuel oil/kerosene, in dollars, 2015

```{r}
#| label: recs-cb-dollarfo
#| echo: FALSE
recs %>%
  summarize(Minimum = min(DOLLARFO, na.rm = TRUE),
            Median = median(DOLLARFO, na.rm = TRUE),
            Maximum = max(DOLLARFO, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTALBTU {-}

Total usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-totalbtu
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTALBTU, na.rm = TRUE),
            Median = median(TOTALBTU, na.rm = TRUE),
            Maximum = max(TOTALBTU, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### TOTALDOL {-}

Total usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-totaldol
#| echo: FALSE
recs %>%
  summarize(Minimum = min(TOTALDOL, na.rm = TRUE),
            Median = median(TOTALDOL, na.rm = TRUE),
            Maximum = max(TOTALDOL, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTUWOOD {-}

Total cordwood usage, in thousand Btu, 2015 (Wood consumption is not included in TOTALBTU or TOTALDOL)

```{r}
#| label: recs-cb-btuwood
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTUWOOD, na.rm = TRUE),
            Median = median(BTUWOOD, na.rm = TRUE),
            Maximum = max(BTUWOOD, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```

#### BTUPELLET {-}

Total wood pellet usage, in thousand Btu, 2015 (Wood consumption is not included in TOTALBTU or TOTALDOL)

```{r}
#| label: recs-cb-btupellet
#| echo: FALSE
recs %>%
  summarize(Minimum = min(BTUPELLET, na.rm = TRUE),
            Median = median(BTUPELLET, na.rm = TRUE),
            Maximum = max(BTUPELLET, na.rm = TRUE)) %>%
  kbl() %>%
  kable_minimal()
```