# RECS Derived Variable Codebook {#recs-cb}

```{r}
#| label: recs-cb-cb-setup
#| echo: FALSE
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(janitor)
library(kableExtra)
library(knitr)

recs <- recs_in %>% haven::zap_labels()
```

The full codebook with the original variables is available at [https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata](https://www.eia.gov/consumption/residential/data/2015/index.php?view=microdata).

## Weighting variables

#### DOEID {-}

Unique identifier for each respondent

#### NWEIGHT {-}

Final sample weight

#### BRRWT1-BRRWT96 {-}

Replicate weights

## Categorical 

#### Region {-}

```{r}
#| label: recs-cb-region
#| echo: FALSE
recs %>%
  cb_count(Region)
```
Variables used: REGIONC

#### Division {-}

```{r}
#| label: recs-cb-division
#| echo: FALSE
recs %>%
  cb_count(Division)
```
Variables used: DIVISION

#### MSAStatus {-}

```{r}
#| label: recs-cb-msastatus
#| echo: FALSE
recs %>%
  cb_count(MSAStatus) 
```
Variables used: METROMICRO

#### Urbanicity {-}

```{r}
#| label: recs-cb-urbanicity
#| echo: FALSE
recs %>%
  cb_count(Urbanicity)
```

Variables used: UATYP10

#### HousingUnitType {-}

```{r}
#| label: recs-cb-housingunittype
#| echo: FALSE
recs %>%
  cb_count(HousingUnitType) 
```

Variables used: TYPEHUQ

#### YearMade {-}

```{r}
#| label: recs-cb-yearmade
#| echo: FALSE
recs %>%
  cb_count(YearMade) 
```

Variables used: YEARMADERANGE

#### SpaceHeatingUsed {-}

```{r}
#| label: recs-cb-spaceheadingused
#| echo: FALSE
recs %>%
  cb_count(SpaceHeatingUsed)
```

Variables used: HEATHOME

#### HeatingBehavior {-}

```{r}
#| label: recs-cb-heatingbehavior
#| echo: FALSE
recs %>%
  cb_count(HeatingBehavior) 
```

Variables used: EQUIPMUSE

#### ACUsed {-}

```{r}
#| label: recs-cb-acused
#| echo: FALSE
recs %>%
  cb_count(ACUsed)
```

Variables used: AIRCOND

#### ACBehavior {-}

```{r}
#| label: recs-cb-acbehavior
#| echo: FALSE
recs %>%
  cb_count(ACBehavior) 
```

Variables used: USECENAC

#### ClimateRegion_BA {-}

```{r}
#| label: recs-cb-climateregion-ba
#| echo: FALSE
recs %>%
  cb_count(ClimateRegion_BA)
```

Variables used: CLIMATE_REGION_PUB

#### ClimateRegion_IECC {-}

```{r}
#| label: recs-cb-climateregion-iecc
#| echo: FALSE
recs %>%
  cb_count(ClimateRegion_IECC)
```

Variables used: IECC_CLIMATE_PUB

## Continuous

#### WinterTempDay {-}

```{r}
#| label: recs-cb-wintertempday
#| echo: FALSE
recs %>%
  cb_continuous(WinterTempDay)
```

Variables used: TEMPHOME

#### WinterTempAway {-}

```{r}
#| label: recs-cb-wintertempaway
#| echo: FALSE
recs %>%
  cb_continuous(WinterTempAway)
```

Variables used: TEMPGONE

#### WinterTempNight {-}

```{r}
#| label: recs-cb-wintertempnight
#| echo: FALSE
recs %>%
  cb_continuous(WinterTempNight)
```

Variables used: TEMPNITE

#### SummerTempDay {-}

```{r}
#| label: recs-cb-summertempday
#| echo: FALSE
recs %>%
  cb_continuous(SummerTempDay)
```

Variables used: TEMPHOMEAC

#### SummerTempAway {-}

```{r}
#| label: recs-cb-summertempaway
#| echo: FALSE
recs %>%
  cb_continuous(SummerTempAway)
```

Variables used: TEMPGONEAC

#### SummerTempNight {-}

```{r}
#| label: recs-cb-summertempnight
#| echo: FALSE
recs %>%
  cb_continuous(SummerTempNight)
```

Variables used: TEMPNITEAC

#### TOTCSQFT {-}

Total cooled square footage

```{r}
#| label: recs-cb-totcsqft
#| echo: FALSE
recs %>%
  cb_continuous(TOTCSQFT)
```

#### TOTHSQFT {-}

Total heated square footage

```{r}
#| label: recs-cb-tothsqft
#| echo: FALSE
recs %>%
  cb_continuous(TOTHSQFT)
```

#### TOTSQFT_EN {-}

Total square footage (used for publication)

```{r}
#| label: recs-cb-totsqft
#| echo: FALSE
recs %>%
  cb_continuous(TOTSQFT_EN)
```

#### TOTUCSQFT {-}

Total uncooled square footage

```{r}
#| label: recs-cb-totucsqft
#| echo: FALSE
recs %>%
  cb_continuous(TOTUCSQFT)
```

#### TOTUSQFT {-}

Total unheated square footage

```{r}
#| label: recs-cb-totusqft
#| echo: FALSE
recs %>%
  cb_continuous(TOTUSQFT)
```

#### CDD30YR {-}

Cooling degree days, 30-year average 1981-2010, base temperature 65F

```{r}
#| label: recs-cb-cdd30yr
#| echo: FALSE
recs %>%
  cb_continuous(CDD30YR)
```

#### CDD65 {-}

Cooling degree days in 2015, base temperature 65F

```{r}
#| label: recs-cb-cdd65
#| echo: FALSE
recs %>%
  cb_continuous(CDD65)
```

#### CDD80 {-}

Cooling degree days in 2015, base temperature 80F (used for garage cooling load estimation only)

```{r}
#| label: recs-cb-cdd80
#| echo: FALSE
recs %>%
  cb_continuous(CDD80)
```

#### HDD30YR {-}

Heating degree days, 30-year average 1981-2010, base temperature 65F

```{r}
#| label: recs-cb-hdd30yr
#| echo: FALSE
recs %>%
  cb_continuous(HDD30YR)
```

#### HDD65 {-}

Heating degree days in 2015, base temperature 65F

```{r}
#| label: recs-cb-hdd65
#| echo: FALSE
recs %>%
  cb_continuous(HDD65)
```

#### HDD50 {-}

Heating degree days in 2015, base temperature 50F (used for garage heating load estimation only)

```{r}
#| label: recs-cb-hdd50
#| echo: FALSE
recs %>%
  cb_continuous(HDD50)
```

#### GNDHDD65 {-}

Heating degree days of ground temperature in 2015, base temperature 65F

```{r}
#| label: recs-cb-gndhdd65
#| echo: FALSE
recs %>%
  cb_continuous(GNDHDD65)
```

#### BTUEL {-}

Total site electricity usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btuel
#| echo: FALSE
recs %>%
  cb_continuous(BTUEL)
```

#### DOLLAREL {-}

Total electricity cost, in dollars, 2015

```{r}
#| label: recs-cb-dollarel
#| echo: FALSE
recs %>%
  cb_continuous(DOLLAREL)
```

#### BTUNG {-}

Total natural gas usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btung
#| echo: FALSE
recs %>%
  cb_continuous(BTUNG)
```

#### DOLLARNG {-}

Total natural gas cost, in dollars, 2015

```{r}
#| label: recs-cb-dollarng
#| echo: FALSE
recs %>%
  cb_continuous(DOLLARNG)
```

#### BTULP {-}

Total propane usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btulp
#| echo: FALSE
recs %>%
  cb_continuous(BTULP)
```

#### DOLLARLP {-}

Total cost of propane, in dollars, 2015

```{r}
#| label: recs-cb-dollarlp
#| echo: FALSE
recs %>%
  cb_continuous(DOLLARLP)
```

#### BTUFO {-}

Total fuel oil/kerosene usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-btufo
#| echo: FALSE
recs %>%
  cb_continuous(BTUFO)
```

#### DOLLARFO {-}

Total cost of fuel oil/kerosene, in dollars, 2015

```{r}
#| label: recs-cb-dollarfo
#| echo: FALSE
recs %>%
  cb_continuous(DOLLARFO)
```

#### TOTALBTU {-}

Total usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-totalbtu
#| echo: FALSE
recs %>%
  cb_continuous(TOTALBTU)
```

#### TOTALDOL {-}

Total usage, in thousand Btu, 2015

```{r}
#| label: recs-cb-totaldol
#| echo: FALSE
recs %>%
  cb_continuous(TOTALDOL)
```

#### BTUWOOD {-}

Total cordwood usage, in thousand Btu, 2015 (Wood consumption is not included in TOTALBTU or TOTALDOL)

```{r}
#| label: recs-cb-btuwood
#| echo: FALSE
recs %>%
  cb_continuous(BTUWOOD)
```

#### BTUPELLET {-}

Total wood pellet usage, in thousand Btu, 2015 (Wood consumption is not included in TOTALBTU or TOTALDOL)

```{r}
#| label: recs-cb-btupellet
#| echo: FALSE
recs %>%
  cb_continuous(BTUPELLET)
```