---
title: "National Crime Victimization Survey (NCVS) 2021 Data Prep"
output: 
  github_document:
    html_preview: false
bibliography: ../book.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data information

Complete data is not stored on this repository but can be obtained on [ICPSR](https://www.icpsr.umich.edu/web/ICPSR/studies/38429) by downloading the R version of data files (@ncvs_data_2021). The files used here are from Version 1 and were downloaded on March 11, 2023.

This script selects a subset of columns of several files and only retains those on this repository.

```{r}
#| label: loadpackageh
#| message: FALSE
library(here) #easy relative paths
```

```{r}
#| label: loadpackages
library(tidyverse) #data manipulation
library(tidylog) #informative logging messages
```

## Incident data file

```{r}
#| label: incfile

incfiles <- load(here("RawData", "NCVS_2021", "ICPSR_38429", "DS0004", "38429-0004-Data.rda"),
     verbose=TRUE)

inc_in <- get(incfiles) %>%
  as_tibble()

inc_slim <- inc_in %>%
  select(
    YEARQ, IDHH, IDPER, V4012, WGTVICCY, # identifiers and weight
    num_range("V", 4016:4019), # series crime information
    V4021B, V4022, V4024, # time of incident, location of incident (macro and micro)
    num_range("V", 4049:4058), #weapon type
    V4234, V4235, num_range("V", 4241:4245), V4248, num_range("V", 4256:4278), starts_with("V4277"), # victim-offender relationship
    V4399, # report to police
    V4529 # type of crime
  )

summary(inc_slim)

write_rds(inc_slim, here("AnalysisData", "ncvs_2021_incident.rds"))
```


## Person data file

```{r}
#| label: persfile

persfiles <- load(here("RawData", "NCVS_2021", "ICPSR_38429", "DS0003", "38429-0003-Data.rda"),
     verbose=TRUE)

pers_in <- get(persfiles) %>%
  as_tibble()

pers_slim <- pers_in %>%
  select(
    YEARQ, IDHH, IDPER, WGTPERCY, # identifiers and weight
    V3014, V3015, V3018, V3023A, V3024, V3084, V3086 
    # age, marital status, sex, race, hispanic origin, gender, sexual orientation
  )

summary(pers_slim)

write_rds(pers_slim, here("AnalysisData", "ncvs_2021_person.rds"))
```

## Household data file


```{r}
#| label: hhfile

hhfiles <- load(here("RawData", "NCVS_2021", "ICPSR_38429", "DS0002", "38429-0002-Data.rda"),
     verbose=TRUE)

hh_in <- get(hhfiles) %>%
  as_tibble()

hh_slim <- hh_in %>%
  select(
    YEARQ, IDHH, WGTHHCY, V2117, V2118, # identifiers, weight, design
    V2015, V2143, SC214A, V2122, V2126B, V2127B, V2129
    # tenure, urbanicity, income, family structure, place size, region, msa status
  )

summary(hh_slim)

write_rds(hh_slim, here("AnalysisData", "ncvs_2021_household.rds"))
```

## Resources

- [USER’S GUIDE TO NATIONAL CRIME VICTIMIZATION SURVEY (NCVS) DIRECT VARIANCE ESTIMATION](https://bjs.ojp.gov/sites/g/files/xyckuh236/files/media/document/ncvs_variance_user_guide_11.06.14.pdf)
-[Appendix C: Examples in SAS](https://bjs.ojp.gov/sites/g/files/xyckuh236/files/media/document/variance_guide_appendix_c_sas.pdf)